### 2017-10-20 paper/draft.tex
Added working draft of project proposal

\documentclass{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{nips_2017}
\usepackage{natbib}
\setcitestyle{numbers,open={[},close={]}}

\title{Deep Multimodal Generative Models for Weakly-Supervised Learning}

\author{
    Mike Wu, Noah Goodman \\
    Department of Computer Science \\
    Stanford University \\
    \texttt{{wumike, ngoodman}@cs.stanford.edu} \\
}

\begin{document}

\maketitle

\begin{abstract}
TODO
\end{abstract}

\section{Introduction}
The problem of learning from multiple modalities is especially important since the majority of modern information is represented through multiple channels. For example, images in the web are often embedded around text. And while images can be described by pixels, we expect related modalities like text to also hold relevant features. People often interact with this multimodal information in a \textit{bi-directional} way. For example, one can imagine what a black cat looks like but also prescribe this description to a photo. Additionally, available multimodal information is often \textit{sparse}. It is sparse in the sense there exist many large corpuses of images or text alone but few datasets exist (and are expensive to construct) that contain multiple modalities together. Ideally, we would like the generative models we build to extract a joint representation that captures high-level concepts across modalities in a way that is bi-directional and weakly-supervised.

In this work, we propose a multimodal variational autoencoder (MMVAE). In our graphical model, we have two modalities $x_{1}$ and $x_{2}$ that are each conditioned on a latent variable $z$, which models the joint distribution, $p(x_{1}, x_{2})$. Such a modal is bi-directional since we can reconstruct both modalities from either $x_{1}$ or $x_{2}$. We will also show that we can train the MMVAE with sparse $(x_{1}, x_{2})$ pairs, and heavily utilize on a larger set of independent $(x_{1})$ and $(x_{2})$ examples. The most significant feature of MMVAEs is that it does not create a separate multimodal encoder $q(z | x_{1}, x_{2})$; instead it shares parameters with the encoders from each modality, $q(z | x_{1})$, $q(z | x_{2})$. This has the benefit of producing a simpler model while ensuring that high-dimensional missing modalities do not cause our generated samples to collapse.

\subsection{Related Work}
In deep learning, multimodal learning is often done using separate branches in the neural net for each modality, and a common top hidden layer. For instance, \citet{ngiam2011multimodal} trained deep autoencoders on audio and video input and found that the bimodal representations were better than any single equivalent. More recently, VAEs (\cite{kingma2013auto, kingma2014semi}) have been used to train models in high-dimensional multimodal settings. For example, conditional VAEs (\cite{sohn2015learning}) maximize a conditional log-likleihood by variational methods where often one modality is conditioned on another i.e. handwriting digits and labels, faces and attributes, or captioning. Notably, conditional VAEs (CVAE) are not bi-directional.

\citet{pandey2017variational} proposed the conditional multimodal autoencoder (CMMA), which also maximizes a conditional log-likelihood. CMMA explicitly connects the conditional variable the latent variable, where CVAE does not. Moreover, it also constrains the latent representation to be close to the joint representation across modalities. However, CMMA is still not bi-directional.

Finally, \citet{suzuki2016joint} proposed a joint multimodal VAE (JMVAE). This is similar to our work in that it is bi-directional, handles missing modalities, and models the joint distribution. Our proposed model (MMVAE) and JMVAE share the same graphical notation, but differ in our approach. Crucially, JMVAE trains separate encoders for each modality and an additional one for the multimodal case. It then includes a KL divergence term \textit{per modality} in the loss. In constract, MMVAE does not have an additional encoder for the multimodal case and does not have additional KL terms. It does so by sharing encoder parameters to learn the guide distributions: $q_{\phi_{x_{1}}}(z|x_{1})$, $q_{\phi_{x_{2}}}(z|x_{2})$, $q_{\phi}(z|x_{1}, x_{2})$

\section{Methods}
This section first briefly goes over VAEs, and product of Gaussians (PoG), then details a new multimodal VAE (MMVAE).

\subsection{Variational Autoencoders}
\subsection{Product of Gaussians}
\subsection{Multimodal Variational Autoencoders}
\subsection{Bidirectional Inference}

\bibliographystyle{abbrvnat}
{\small
\linespread{1}
\bibliography{draft}
}

\end{document}

### 2017-10-20 paper/draft.tex
added some background methods -- need to add description for MMVAE

@@
 \usepackage{nips_2017}
+\usepackage{graphicx}
 \usepackage{natbib}
+
 \setcitestyle{numbers,open={[},close={]}}
+\graphicspath{ {./images/} }
 
@@
 For instance, \citet{ngiam2011multimodal} trained deep autoencoders on audio and video input and found that the bimodal representations were better than any single equivalent.
-More recently, VAEs (\cite{kingma2013auto, kingma2014semi}) have been used to train models in high-dimensional multimodal settings.
-For example, conditional VAEs (\cite{sohn2015learning}) maximize a conditional log-likleihood by variational methods where often one modality is conditioned on another i.e.
+More recently, VAEs \cite{kingma2013auto, kingma2014semi} have been used to train models in high-dimensional multimodal settings.
+For example, conditional VAEs \cite{sohn2015learning} maximize a conditional log-likleihood by variational methods where often one modality is conditioned on another i.e.
 handwriting digits and labels, faces and attributes, or captioning.
@@
 \subsection{Variational Autoencoders}
-\subsection{Product of Gaussians}
+Variational autoencoders (VAE) model a generative process as $z \sim \mathcal{N}(0, I)$ and $x \sim p_{\theta}(x | z)$ where $x$ is an observed variable, $z$ is a latent variable, and $\theta$ are the model parameters.
+VAE wish to maximize the marginal distribution $p(x) = \int p_{\theta}(x | z)p(z)dx$, which is intractor.
+Instead, we maximize the lower bound of the marginal distribution:
+
+\begin{equation}
+    \textup{log } p(x) \geq -D_{KL}(q_{\phi}(z | x) || p(z)) + E_{q_{\phi}}[\textup{log } p_{\theta}(x | z)]
+\end{equation}
+
+where $q_{\theta}(z | x)$ is a \textit{guide distribution} that approximates the posterior $p(z | x)$; $\theta$ are the parameters of $q$.
+In the text, we call $p_{\theta}(x | z)$ a decoder and $q_{\theta}(z | x)$ an encoder.
+In practice, $q$ is set to the familiy of Gaussian distributions, where optimization amounts to learning $\phi=\{\mu, \sigma\}$.
+The standard reparametrization trick is employed to sample $z$.
+
+\subsection{Product of Experts}
+
+We wish to model the interaction between our modalities as a product of experts.
+Fortunately, because our guide distributions are Gaussian, MMVAEs take advantage of the fact the product of two Gaussian distributions is itself a Gaussian distribution \cite{cao2014generalized}.
+The mean and covariance of the resulting Gaussian distribution is:
+
+\begin{equation}
+    \mu = (\sum_{i} \mu_{i}\Sigma^{-1}_{i})(\sum_{i}\Sigma^{-1}_{i})^{-1}
+\end{equation}
+\begin{equation}
+    \Sigma = (\sum_{i} \Sigma^{-1}_{i})^{-1}
+\end{equation}
+
+where $\mu_{i}$, $\Sigma_{i}$ is the mean and covariance of the $i$-th Gaussian expert.
+
 \subsection{Multimodal Variational Autoencoders}
 \subsection{Bidirectional Inference}
+
+\section{Experiments}
+TODO
 

### 2017-10-21 paper/draft.tex
Added more text

@@
 \usepackage{graphicx}
+\usepackage{subcaption}
+\usepackage{amsmath}
 \usepackage{natbib}
@@
 \subsection{Multimodal Variational Autoencoders}
-\subsection{Bidirectional Inference}
+Given a dataset $(X, W) = \{ (x_{1,1}, x_{2,1}), ..., (x_{1,N}, x_{2,N}) \}$, where $x_{1}$ and $x_{2}$ are two modalities of possibly different dimensions.
+We assume that $x_{1}$ and $x_{2}$ are conditionally independent on the latent state $z$.
+Figure \ref{fig:diagram:graph} shows the graphical model of the generative process: $z \sim p(z)$ and $x_{1}, x_{2} \sim p(x_{1}, x_{2} | z)$.
 
-\section{Experiments}
-TODO
+We can estimate the lower bound of the log-likelihood $\textup{log }p(x_{1}, x_{2})$ as follows:
+
+\begin{equation}
+\textup{log }p(x_{1}, x_{2}) \geq E_{q_{\theta}(z | x_{1}, x_{2})}[\textup{log }\frac{p_{\theta}(x_{1}, x_{2}, z)}{q_{\phi}(z | x_{1}, x_{2})}]
+\end{equation}
+
+\begin{multline}
+\textup{log }p(x_{1}, x_{2}) \geq -D_{KL}(q_{\phi}(z | x_{1}, x_{2}) || p(z)) \\ + E_{q_{\phi}(z | x_{1}, x_{2})}[\textup{log } p_{\theta_{x_{1}}}(x_{1} | z)] + E_{q_{\phi}(z | x_{1}, x_{2})}[\textup{log } p_{\theta_{x_{2}}}(x_{2} | z)]
+\end{multline}
+
+The equation above has two reconstruction terms, one for each modality.
+As with VAEs, we have $q_{\theta}(z | x_{1}, x_{2})$.
+as the encoder and $p_{\theta_{x_{1}}}(x_{1} | z)$ and $p_{\theta_{x_{2}}}(x_{2} | z)$ as decoders.
+
+\begin{figure}[!h]
+\centering
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=.75\linewidth]{graph.pdf}
+        \caption{}
+        \label{fig:diagram:graph}
+    \end{subfigure}\hspace{5mm}
+    \begin{subfigure}[b]{.10\linewidth}
+        \includegraphics[width=.75\linewidth]{modelv2}
+        \caption{}
+        \label{fig:diagram:modelv2}
+    \end{subfigure}\hspace{5mm}
+    \begin{subfigure}[b]{.10\linewidth}
+        \includegraphics[width=.75\linewidth]{modelv1}
+        \caption{}
+        \label{fig:diagram:modelv1}
+    \end{subfigure}\hspace{5mm}
+    \begin{subfigure}[b]{.15\linewidth}
+        \includegraphics[width=.75\linewidth]{modelv3}
+        \caption{}
+        \label{fig:diagram:modelv3}
+    \end{subfigure}
+    \caption{(a) Graphical model of the MMVAE.
+Gray circles represent observed variables.
+The white circles represent latent variables.
+(b) MMVAE when modality 2 is missing.
+When only 1 modality is present, we ignore the PoG (product of Gaussians).
+(c) MMVAE when modality 1 is missing.
+(d) MMVAE when both modalities exist.
+$E_{1}$ and $E_{2}$ represent the two encoders: $q_{\phi_{x_{1}}}(z | x_{1})$ and $q_{\phi_{x_{2}}}(z | x_{2})$.
+Regardless if modalities are missing, we optimize the parameters of $E_{1}$ and $E_{2}$.}
+    \label{fig:diagram}
+\end{figure}
 

### 2017-10-21 paper/draft.tex
Edits before turning in the proposal

@@
 For example, one can imagine what a black cat looks like but also prescribe this description to a photo.
-Additionally, available multimodal information is often \textit{sparse}.
-It is sparse in the sense there exist many large corpuses of images or text alone but few datasets exist (and are expensive to construct) that contain multiple modalities together.
+Additionally, available multimodal information is often \textit{sparse}, in the sense there exist many large corpuses of images or text alone but few datasets exist (and are expensive to construct) that contain multiple modalities together.
 Ideally, we would like the generative models we build to extract a joint representation that captures high-level concepts across modalities in a way that is bi-directional and weakly-supervised.
@@
 It then includes a KL divergence term \textit{per modality} in the loss.
-In constract, MMVAE does not have an additional encoder for the multimodal case and does not have additional KL terms.
+In constrast, MMVAE does not have an additional encoder for the multimodal case and does not have additional KL terms.
 It does so by sharing encoder parameters to learn the guide distributions: $q_{\phi_{x_{1}}}(z|x_{1})$, $q_{\phi_{x_{2}}}(z|x_{2})$, $q_{\phi}(z|x_{1}, x_{2})$
@@
 Variational autoencoders (VAE) model a generative process as $z \sim \mathcal{N}(0, I)$ and $x \sim p_{\theta}(x | z)$ where $x$ is an observed variable, $z$ is a latent variable, and $\theta$ are the model parameters.
-VAE wish to maximize the marginal distribution $p(x) = \int p_{\theta}(x | z)p(z)dx$, which is intractor.
+VAE wish to maximize the marginal distribution $p(x) = \int p_{\theta}(x | z)p(z)dx$, which is intractable.
 Instead, we maximize the lower bound of the marginal distribution:
@@
 
-where $q_{\theta}(z | x)$ is a \textit{guide distribution} that approximates the posterior $p(z | x)$; $\theta$ are the parameters of $q$.
+where $q_{\phi}(z | x)$ is a \textit{guide distribution} that approximates the posterior $p(z | x)$; $\phi$ are the parameters of $q$.
 In the text, we call $p_{\theta}(x | z)$ a decoder and $q_{\theta}(z | x)$ an encoder.
@@
 \subsection{Multimodal Variational Autoencoders}
-Given a dataset $(X, W) = \{ (x_{1,1}, x_{2,1}), ..., (x_{1,N}, x_{2,N}) \}$, where $x_{1}$ and $x_{2}$ are two modalities of possibly different dimensions.
+Given a dataset $(X_{1}, X_{2}) = \{ (x_{1,1}, x_{2,1}), ..., (x_{1,N}, x_{2,N}) \}$, where $x_{1}$ and $x_{2}$ are two modalities of possibly different dimensions.
 We assume that $x_{1}$ and $x_{2}$ are conditionally independent on the latent state $z$.
@@
 We can estimate the lower bound of the log-likelihood $\textup{log }p(x_{1}, x_{2})$ as follows:
-
-\begin{equation}
-\textup{log }p(x_{1}, x_{2}) \geq E_{q_{\theta}(z | x_{1}, x_{2})}[\textup{log }\frac{p_{\theta}(x_{1}, x_{2}, z)}{q_{\phi}(z | x_{1}, x_{2})}]
-\end{equation}
 
@@
 The equation above has two reconstruction terms, one for each modality.
-As with VAEs, we have $q_{\theta}(z | x_{1}, x_{2})$.
-as the encoder and $p_{\theta_{x_{1}}}(x_{1} | z)$ and $p_{\theta_{x_{2}}}(x_{2} | z)$ as decoders.
+As with VAEs, we can parametrize the encoder and decoder as deep neural networks and use stochastic gradient descent to optimize the parameters, $\theta_{x_{1}}$, $\theta_{x_{2}}$, $\phi$.
+Because each modality learns a different feature representation, each modality has its own encoder and decoder.
 
-\begin{figure}[!h]
+In MMVAE, it is clear that we can compute the marginal and conditional distribution in both directions: given either $x_{1}$ or $x_{2}$ we can reconstruct both modalities.
+Additionally, we can easily extend MMVAE to handle more than two modalities, $p(x_{1}, x_{2}, ..., x_{n})$ using the same framework.
+
+\begin{figure}[h!]
 \centering
@@
 
+\subsection{Weakly-Supervised Training}
+In most circumstances, paired modal data is both expensive and scarce while unimodal data, i.e.
+images or text is more easily available.
+The design of MMVAE allows for bootstrapping under sparse labels.
+During training, we consider input pairs of 3 types: the 1st modality alone,  the 2nd modality alone, or two separate modalities.
+Figure \ref{fig:diagram:modelv2}, \ref{fig:diagram:modelv1}, \ref{fig:diagram:modelv3} enumerate these three cases.
+In the case of a single modality, training is identical to VAE; the encoder $E_{i}(x_{i})$ returns $\mu_{i}$, $\sigma_{i}$, which is used to sample $z$ and decode.
+In the case of two modalities, each encoder returns $\mu_{i}$, $\sigma_{i}$, which we can combine via PoG to return the parameters of the joint Gaussian distribution.
+This is then used to sample $z$ and decode as in the traditional VAE.
+Crucially, we do not need additional loss terms or separate networks to handle unimodal and multimodal cases; the only change required is an additional PoG operation.
+In practice, we find that we only need a limited amount of paired multimodal examples in order to learn the cross-modal relations.
+
 \bibliographystyle{abbrvnat}

### 2017-10-21 mnist/train.py
Added multimnist file for generating dataset

@@ -89,11 +89,11 @@ if __name__ == "__main__":
     # create loaders for MNIST
     train_loader = torch.utils.data.DataLoader(
         datasets.MNIST('./data', train=True, download=True,
-                     transform=transforms.ToTensor()),
+                       transform=transforms.ToTensor()),
         batch_size=args.batch_size, shuffle=True)
     test_loader = torch.utils.data.DataLoader(
         datasets.MNIST('./data', train=False, download=True,
-                     transform=transforms.ToTensor()),
+                       transform=transforms.ToTensor()),
         batch_size=args.batch_size, shuffle=True)
 
     # load multimodal VAE

### 2017-10-21 multimnist/dataset.py

@@ -1,0 +1,106 @@
+"""
+This script generates a dataset similar to the Multi-MNIST dataset
+described in [1].
+
+[1] Eslami, SM Ali, et al. "Attend, infer, repeat: Fast scene
+understanding with generative models." Advances in Neural Information
+Processing Systems. 2016.
+"""
+
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import os
+import numpy as np
+import torch
+import torchvision.datasets as dset
+from scipy.misc import imresize
+
+
+def sample_one(canvas_size, mnist):
+    i = np.random.randint(mnist['digits'].shape[0])
+    digit = mnist['digits'][i]
+    label = mnist['labels'][i]
+    scale = 0.1 * np.random.randn() + 1.3
+    resized = imresize(digit, 1. / scale)
+    w = resized.shape[0]
+    assert w == resized.shape[1]
+    padding = canvas_size - w
+    pad_l = np.random.randint(0, padding)
+    pad_r = np.random.randint(0, padding)
+    pad_width = ((pad_l, padding - pad_l), (pad_r, padding - pad_r))
+    positioned = np.pad(resized, pad_width, 'constant', constant_values=0)
+    return positioned, label
+
+
+def sample_multi(num_digits, canvas_size, mnist):
+    canvas = np.zeros((canvas_size, canvas_size))
+    labels = []
+    for _ in range(num_digits):
+        positioned_digit, label = sample_one(canvas_size, mnist)
+        canvas += positioned_digit
+        labels.append(label)
+    
+    # Crude check for overlapping digits.
+    if np.max(canvas) > 255:
+        return sample_multi(num_digits, canvas_size, mnist)
+    else:
+        return canvas, labels
+
+
+def mk_dataset(n, mnist, min_digits, max_digits, canvas_size):
+    x = []
+    y = []
+    for _ in range(n):
+        num_digits = np.random.randint(min_digits, max_digits + 1)
+        canvas, labels = sample_multi(num_digits, canvas_size, mnist)
+        x.append(canvas)
+        y.append(labels)
+    return np.array(x, dtype=np.uint8), y
+
+
+def load_mnist():
+    train_loader = torch.utils.data.DataLoader(
+        dset.MNIST(root='./data', train=True, download=True))
+
+    test_loader = torch.utils.data.DataLoader(
+        dset.MNIST(root='./data', train=False, download=True))
+    
+    train_data = {
+        'digits': train_loader.dataset.train_data.numpy(),
[…]

### 2017-10-21 mnist/test.py
Added script to measure MNIST performance of MMVAE

@@ -1,0 +1,57 @@
+"""Test the MMVAE trained on MNIST on the MNIST task: 
+Given the test set of images, generate the text and see if 
+it is the correct label. Report the accuracy score.
+"""
+
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import torch
+import torch.nn.functional as F
+from torch.autograd import Variable
+
+from train import load_checkpoint
+
+
+def test_mnist(model, loader):
+    model.eval()
+    test_loss = 0
+    correct = 0
+    for data, target in loader:
+        if args.cuda:
+            data, target = data.cuda(), target.cuda()
+        data, target = Variable(data, volatile=True), Variable(target)
+        output = model(data)
+        test_loss += F.nll_loss(output, target, size_average=False).data[0]
+        pred = output.data.max(1, keepdim=True)[1]
+        correct += pred.eq(target.data.view_as(pred)).cpu().sum()
+
+    test_loss /= len(loader.dataset)
+    print('\nTest set: Average loss: {:.4f}, Accuracy: {}/{} ({:.0f}%)\n'.format(
+        test_loss, correct, len(loader.dataset),
+        100. * correct / len(loader.dataset)))
+
+    return correct / float(len(loader.dataset))
+
+
+if __name__ == "__main__":
+    import argparse
+    parser = argparse.ArgumentParser()
+    parser.add_argument('model_path', type=str, help='path to trained model file')
+    parser.add_argument('--cuda', action='store_true', default=False,
+                        help='enables CUDA training')
+    args = parser.parse_args()
+    args.cuda = args.cuda and torch.cuda.is_available()
+
+
+    # loader for MNIST
+    loader = torch.utils.data.DataLoader(
+        datasets.MNIST('./data', train=False, download=True,
+                       transform=transforms.ToTensor()),
+        batch_size=128, shuffle=True)
+
+    vae = load_checkpoint(args.model_path, use_cuda=args.cuda)
+    vae.eval()
+
+    test_mnist(vae, loader)

### 2017-10-21 mnist/train.py

@@ -52,6 +52,9 @@ def load_checkpoint(file_path, n_latents=20, use_cuda=False):
     vae = MultimodalVAE(n_latents=n_latents)
     vae.load_state_dict(checkpoint['state_dict'])
     
+    if use_cuda:
+        vae.cuda()
+
     return vae
 
 

### 2017-10-21 mnist/train.py
Added weak supervision test to train a bunch of models at different rates of weak supervision

@@ -134,7 +134,7 @@ if __name__ == "__main__":
             
             loss = loss_function(recon_image, image_3x, recon_text, text_3x, mu, logvar)
             loss.backward()
-            loss_meter.update(loss.data[0], len(image))
+            loss_meter.update(loss.data[0], len(image) * 3.)  # we show image 3 times 
             optimizer.step()
 
             if batch_idx % args.log_interval == 0:

### 2017-10-21 mnist/weak.py

@@ -1,0 +1,190 @@
+"""Weak supervision test. This will loop through and train
+a number of MMVAE models where we progressively cut the 
+number of examples we show it in half. The point is to test 
+how well the model is able to learn on sparse relation data.
+"""
+
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import os
+import sys
+import shutil
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.autograd import Variable
+from torchvision import transforms, datasets
+
+from model import MultimodalVAE
+from train import AverageMeter
+from train import save_checkpoint, load_checkpoint
+from train import loss_function
+
+
+def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, lr=1e-3, 
+                   log_interval=10, cuda=False):
+    """Pipeline to train and test MultimodalVAE on MNIST dataset. This is 
+    identical to the code in train.py.
+
+    :param out_dir: directory to store trained models
+    :param weak_perc: percent of time to show a relation pair (vs no relation pair)
+    :param n_latents: size of latent variable (default: 20)
+    :param batch_size: number of examples to show at once (default: 128)
+    :param epochs: number of loops over dataset (default: 20)
+    :param lr: learning rate (default: 1e-3)
+    :param log_interval: interval of printing (default: 10)
+    :param cuda: whether to use cuda or not (default: False)
+    """
+    # create loaders for MNIST
+    train_loader = torch.utils.data.DataLoader(
+        datasets.MNIST('./data', train=True, download=True,
+                       transform=transforms.ToTensor()),
+        batch_size=batch_size, shuffle=True)
+    test_loader = torch.utils.data.DataLoader(
+        datasets.MNIST('./data', train=False, download=True,
+                       transform=transforms.ToTensor()),
+        batch_size=batch_size, shuffle=True)
+
+    # load multimodal VAE
+    vae = MultimodalVAE(n_latents=n_latents)
+    if cuda:
+        vae.cuda()
+
+    optimizer = optim.Adam(vae.parameters(), lr=lr)
+
+
+    def train(epoch):
+        np.random.seed(42)  # important to have the same seed
+                            # in order to make the same choices for weak supervision
+                            # otherwise, we end up showing different examples over epochs
[…]

### 2017-10-21 mnist/weak.py
Fixed random seed so that we show the same examples over epochs

@@ -11,6 +11,8 @@ from __future__ import absolute_import
 import os
 import sys
 import shutil
+import random
+import numpy as np
 
 import torch
 import torch.nn as nn
@@ -57,7 +59,21 @@ def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, 
     optimizer = optim.Adam(vae.parameters(), lr=lr)
 
 
+    def loss_function(recon_image, image, recon_text, text, mu, logvar):
+        image_BCE = F.binary_cross_entropy(recon_image, image.view(-1, 784))
+        text_BCE = F.nll_loss(recon_text, text)
+
+        # see Appendix B from VAE paper:
+        # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
+        # https://arxiv.org/abs/1312.6114
+        # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
+        KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
+        KLD /= args.batch_size * 784
+        return image_BCE + text_BCE + KLD
+    
+
     def train(epoch):
+        random.seed(42)
         np.random.seed(42)  # important to have the same seed
                             # in order to make the same choices for weak supervision
                             # otherwise, we end up showing different examples over epochs
@@ -86,7 +102,8 @@ def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, 
             n = 2
 
             # if we flip(weak_perc), then we show a paired relation example.
-            if np.random.random() < weak_perc:
+            flip = np.random.random()
+            if flip < weak_perc:
                 recon_image_1, recon_text_1, mu_1, logvar_1 = vae(image, text)
                 recon_image = torch.cat((recon_image, recon_image_1))
                 recon_text = torch.cat((recon_text, recon_text_1))
@@ -174,17 +191,17 @@ if __name__ == "__main__":
     supervision_dir = './trained_models/weak_supervision'
     if os.path.isdir(supervision_dir):
         shutil.rmtree(supervision_dir)
-        os.makedirs(supervision_dir)
-        print('Created directory: %s' % supervision_dir)
+    os.makedirs(supervision_dir)
+    print('Created directory: %s' % supervision_dir)
 
     for weak_perc in [0, 1, 2, 5, 10, 25, 50, 100]:
[…]

### 2017-10-21 mnist/weak.py
bugfixes with weak script

@@ -194,8 +194,8 @@ if __name__ == "__main__":
     os.makedirs(supervision_dir)
     print('Created directory: %s' % supervision_dir)
 
-    for weak_perc in [0, 1, 2, 5, 10, 25, 50, 100]:
-        perc_dir = os.path.join(supervision_dir, 'weak_perc_%d' % weak_perc)
+    for weak_perc in [0, 0.01, 0.02, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.]:
+        perc_dir = os.path.join(supervision_dir, 'weak_perc_{}'.format(weak_perc))
         if os.path.isdir(perc_dir):
             shutil.rmtree(perc_dir)
         os.makedirs(perc_dir)

### 2017-10-22 mnist/test.py
Added script to make plot of performance from each trained weak supervision model

@@ -15,7 +15,7 @@ from torchvision import datasets, transforms
 from train import load_checkpoint
 
 
-def test_mnist(model, loader):
+def test_mnist(model, loader, verbose=True):
     """Functionalize this so we can call it from our weakly-supervised experiments."""
     model.eval()
     correct = 0
@@ -29,8 +29,9 @@ def test_mnist(model, loader):
         pred = recon_text.data.max(1, keepdim=True)[1]
         correct += pred.eq(text.data.view_as(pred)).cpu().sum()
 
-    print('\nTest set: Accuracy: {}/{} ({:.0f}%)\n'.format(
-        correct, len(loader.dataset), 100. * correct / len(loader.dataset)))
+    if verbose:
+        print('\nTest set: Accuracy: {}/{} ({:.0f}%)\n'.format(
+            correct, len(loader.dataset), 100. * correct / len(loader.dataset)))
 
     return correct / float(len(loader.dataset))
 

### 2017-10-22 mnist/test_weak.py

@@ -1,0 +1,50 @@
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import torch
+from torchvision import datasets, transforms
+
+from test import test_mnist
+from train import load_checkpoint
+
+
+if __name__ == "__main__":
+    import argparse
+    from glob import glob
+
+    import matplotlib
+    matplotlib.use('Agg')
+    import matplotlib.pyplot as plt
+
+    parser = argparse.ArgumentParser()
+    parser.add_argument('models_dir', type=str, help='path to output directory of weak.py')
+    parser.add_argument('--cuda', action='store_true', default=False,
+                        help='enables CUDA training')
+    args = parser.parse_args()
+    args.cuda = args.cuda and torch.cuda.is_available()
+
+    x, y = [], []
+
+    for dir_path in glob(os.path.join(args.models_dir, '*')):
+        weak_perc = float(os.path.basename(dir_path).split('_')[-1])
+        loader = torch.utils.data.DataLoader(
+            datasets.MNIST('./data', train=False, download=True,
+                           transform=transforms.ToTensor()),
+            batch_size=128, shuffle=True)
+        vae = load_checkpoint(args.model_path, use_cuda=args.cuda)
+        vae.eval()
+        weak_acc = test_mnist(vae, loader, verbose=False)
+
+        x.append(weak_perc)
+        y.append(weak_acc)
+
+
+    plt.figure()
+    plt.plot(x, y, 'o')
+    plt.xlabel('% Supervision', fontsize=18)
+    plt.ylabel('MNIST Accuracy', fontsize=18)
+    plot.tick_params(axis='both', which='major', labelsize=16)
+    plot.tick_params(axis='both', which='minor', labelsize=16)
+    plt.tight_layout()
+    plt.savefig('./weak_supervision.png')

### 2017-10-22 mnist/test.py
Bugfixes to test script

@@ -15,12 +15,12 @@ from torchvision import datasets, transforms
 from train import load_checkpoint
 
 
-def test_mnist(model, loader, verbose=True):
+def test_mnist(model, loader, use_cuda=False, verbose=True):
     """Functionalize this so we can call it from our weakly-supervised experiments."""
     model.eval()
     correct = 0
     for image, text in loader:
-        if args.cuda:
+        if use_cuda:
             image, text = image.cuda(), text.cuda()
         image, text = Variable(image, volatile=True), Variable(text)
         image = image.view(-1, 784)
@@ -55,4 +55,4 @@ if __name__ == "__main__":
     vae = load_checkpoint(args.model_path, use_cuda=args.cuda)
     vae.eval()
 
-    test_mnist(vae, loader)
+    test_mnist(vae, loader, use_cuda=args.cuda)

### 2017-10-22 mnist/test_weak.py

@@ -10,6 +10,7 @@ from train import load_checkpoint
 
 
 if __name__ == "__main__":
+    import os
     import argparse
     from glob import glob
 
@@ -32,19 +33,20 @@ if __name__ == "__main__":
             datasets.MNIST('./data', train=False, download=True,
                            transform=transforms.ToTensor()),
             batch_size=128, shuffle=True)
-        vae = load_checkpoint(args.model_path, use_cuda=args.cuda)
+        vae = load_checkpoint(os.path.join(dir_path, 'model_best.pth.tar'), use_cuda=args.cuda)
         vae.eval()
-        weak_acc = test_mnist(vae, loader, verbose=False)
+        weak_acc = test_mnist(vae, loader, use_cuda=args.cuda, verbose=False)
 
         x.append(weak_perc)
         y.append(weak_acc)
+        print('Got accuracies for %s.' % dir_path)
 
 
     plt.figure()
-    plt.plot(x, y, 'o')
+    plt.plot(x, y, '-o')
     plt.xlabel('% Supervision', fontsize=18)
     plt.ylabel('MNIST Accuracy', fontsize=18)
-    plot.tick_params(axis='both', which='major', labelsize=16)
-    plot.tick_params(axis='both', which='minor', labelsize=16)
+    plt.tick_params(axis='both', which='major', labelsize=16)
+    plt.tick_params(axis='both', which='minor', labelsize=16)
     plt.tight_layout()
     plt.savefig('./weak_supervision.png')

### 2017-10-22 mnist/weak.py

@@ -112,7 +112,8 @@ def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, 
                 image_nx = torch.cat((image_nx, image))
                 text_nx = torch.cat((text_nx, text))
                 n = 3
-            
+                import pdb; pdb.set_trace()
+
             loss = loss_function(recon_image, image_nx, recon_text, text_nx, mu, logvar)
             loss.backward()
             loss_meter.update(loss.data[0], len(image) * n)

### 2017-10-22 mnist/weak.py
Sparse set

@@ -195,7 +195,7 @@ if __name__ == "__main__":
     os.makedirs(supervision_dir)
     print('Created directory: %s' % supervision_dir)
 
-    for weak_perc in [0, 0.01, 0.02, 0.05, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.]:
+    for weak_perc in [0, 0.01, 0.02, 0.05, 0.1, 0.25, 0.5, 0.75, 1.]:
         perc_dir = os.path.join(supervision_dir, 'weak_perc_{}'.format(weak_perc))
         if os.path.isdir(perc_dir):
             shutil.rmtree(perc_dir)

### 2017-10-22 mnist/datasets.py
Added a ShuffleNIST dataset

@@ -1,0 +1,37 @@
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import torch
+from torch.utils.data.dataset import Dataset
+
+import numpy as np
+from PIL import Image
+
+
+class ShuffleMNIST(Dataset):
+    def __init__(self, data_path, train=True, transform=None):
+        data_path = os.path.join(data_path, 'processed',
+                                 'training.pt' if train else 'test.pt')
+        data, targets = torch.load(data_path)
+        shuffled = targets.numpy()
+        np.random.shuffle(shuffled)
+        shuffled = torch.from_numpy(shuffled)
+        self.data = data
+        self.targets = targets
+        self.shuffled = shuffled
+        self.transform = transform
+
+    def __getitem__(self, index):
+        data = self.data[index]
+        targets = self.targets[index]
+        shuffled = self.shuffled[index]
+
+        if self.transform is not None:
+            data = Image.fromarray(data.numpy(), mode='L')
+            data = self.tranform(data)
+
+        return data, targets, shuffled
+
+    def __len__(self):
+        return len(self.data)

### 2017-10-22 mnist/train.py

@@ -61,7 +61,6 @@ def load_checkpoint(file_path, n_latents=20, use_cuda=False):
 def loss_function(recon_image, image, recon_text, text, mu, logvar):
     image_BCE = F.binary_cross_entropy(recon_image, image.view(-1, 784))
     text_BCE = F.nll_loss(recon_text, text)
-
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
     # https://arxiv.org/abs/1312.6114
@@ -69,6 +68,28 @@ def loss_function(recon_image, image, recon_text, text, mu, logvar):
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
     KLD /= args.batch_size * 784
     return image_BCE + text_BCE + KLD
+
+
+def image_loss_function(recon_image, image, mu, logvar):
+    image_BCE = F.binary_cross_entropy(recon_image, image.view(-1, 784))
+    # see Appendix B from VAE paper:
+    # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
+    # https://arxiv.org/abs/1312.6114
+    # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
+    KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
+    KLD /= args.batch_size * 784
+    return image_BCE + KLD
+
+
+def text_loss_function(recon_text, text, mu, logvar):
+    text_BCE = F.nll_loss(recon_text, text)
+    # see Appendix B from VAE paper:
+    # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
+    # https://arxiv.org/abs/1312.6114
+    # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
+    KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
+    KLD /= args.batch_size * 784
+    return image_BCE + KLD
 
 
 if __name__ == "__main__":

### 2017-10-22 mnist/datasets.py
Fixed problem where i was computing joint loss from single modalities = which is implicitly applying pairs

@@ -1,37 +1,0 @@
-from __future__ import division
-from __future__ import print_function
-from __future__ import absolute_import
-
-import torch
-from torch.utils.data.dataset import Dataset
-
-import numpy as np
-from PIL import Image
-
-
-class ShuffleMNIST(Dataset):
-    def __init__(self, data_path, train=True, transform=None):
-        data_path = os.path.join(data_path, 'processed',
-                                 'training.pt' if train else 'test.pt')
-        data, targets = torch.load(data_path)
-        shuffled = targets.numpy()
-        np.random.shuffle(shuffled)
-        shuffled = torch.from_numpy(shuffled)
-        self.data = data
-        self.targets = targets
-        self.shuffled = shuffled
-        self.transform = transform
-
-    def __getitem__(self, index):
-        data = self.data[index]
-        targets = self.targets[index]
-        shuffled = self.shuffled[index]
-
-        if self.transform is not None:
-            data = Image.fromarray(data.numpy(), mode='L')
-            data = self.tranform(data)
-
-        return data, targets, shuffled
-
-    def __len__(self):
-        return len(self.data)

### 2017-10-22 mnist/train.py

@@ -11,7 +11,7 @@ import torch.nn as nn
 import torch.optim as optim
 import torch.nn.functional as F
 from torch.autograd import Variable
-from torchvision import transforms, datasets
+from torchvision import transforms
 
 from model import MultimodalVAE
 
@@ -58,7 +58,7 @@ def load_checkpoint(file_path, n_latents=20, use_cuda=False):
     return vae
 
 
-def loss_function(recon_image, image, recon_text, text, mu, logvar):
+def joint_loss_function(recon_image, image, recon_text, text, mu, logvar, batch_size=128):
     image_BCE = F.binary_cross_entropy(recon_image, image.view(-1, 784))
     text_BCE = F.nll_loss(recon_text, text)
     # see Appendix B from VAE paper:
@@ -66,29 +66,29 @@ def loss_function(recon_image, image, recon_text, text, mu, logvar):
     # https://arxiv.org/abs/1312.6114
     # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-    KLD /= args.batch_size * 784
+    KLD /= batch_size * 784
     return image_BCE + text_BCE + KLD
 
 
-def image_loss_function(recon_image, image, mu, logvar):
+def image_loss_function(recon_image, image, mu, logvar, batch_size=128):
     image_BCE = F.binary_cross_entropy(recon_image, image.view(-1, 784))
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
     # https://arxiv.org/abs/1312.6114
     # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-    KLD /= args.batch_size * 784
+    KLD /= batch_size * 784
     return image_BCE + KLD
 
 
-def text_loss_function(recon_text, text, mu, logvar):
+def text_loss_function(recon_text, text, mu, logvar, batch_size=128):
     text_BCE = F.nll_loss(recon_text, text)
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
     # https://arxiv.org/abs/1312.6114
[…]

### 2017-10-22 mnist/weak.py

@@ -24,7 +24,7 @@ from torchvision import transforms, datasets
 from model import MultimodalVAE
 from train import AverageMeter
 from train import save_checkpoint, load_checkpoint
-from train import loss_function
+from train import joint_loss_function, image_loss_function, text_loss_function
 
 
 def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, lr=1e-3, 
@@ -59,26 +59,16 @@ def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, 
     optimizer = optim.Adam(vae.parameters(), lr=lr)
 
 
-    def loss_function(recon_image, image, recon_text, text, mu, logvar):
-        image_BCE = F.binary_cross_entropy(recon_image, image.view(-1, 784))
-        text_BCE = F.nll_loss(recon_text, text)
-
-        # see Appendix B from VAE paper:
-        # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
-        # https://arxiv.org/abs/1312.6114
-        # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
-        KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-        KLD /= args.batch_size * 784
-        return image_BCE + text_BCE + KLD
-    
-
     def train(epoch):
         random.seed(42)
         np.random.seed(42)  # important to have the same seed
                             # in order to make the same choices for weak supervision
                             # otherwise, we end up showing different examples over epochs
         vae.train()
-        loss_meter = AverageMeter()
+
+        joint_loss_meter = AverageMeter()
+        image_loss_meter = AverageMeter()
+        text_loss_meter = AverageMeter()
 
         for batch_idx, (image, text) in enumerate(train_loader):
             if cuda:
@@ -89,48 +79,44 @@ def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, 
             
             # for each batch, use 3 types of examples (joint, image-only, and text-only)
             # this way, we can hope to reconstruct both modalities from one
-            recon_image_2, recon_text_2, mu_2, logvar_2 = vae(image=image)
-            recon_image_3, recon_text_3, mu_3, logvar_3 = vae(text=text)
-            # combine all of the batches (no need to reorder; we show the model all at once)
-            recon_image = torch.cat((recon_image_2, recon_image_3))
[…]

### 2017-10-22 mnist/train.py
Small bug fixes

@@ -11,7 +11,7 @@ import torch.nn as nn
 import torch.optim as optim
 import torch.nn.functional as F
 from torch.autograd import Variable
-from torchvision import transforms
+from torchvision import datasets, transforms
 
 from model import MultimodalVAE
 
@@ -89,7 +89,7 @@ def text_loss_function(recon_text, text, mu, logvar, batch_size=128):
     # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
     KLD /= batch_size * 784
-    return image_BCE + KLD
+    return text_BCE + KLD
 
 
 if __name__ == "__main__":
@@ -166,7 +166,7 @@ if __name__ == "__main__":
                     100. * batch_idx / len(train_loader), joint_loss_meter.avg,
                     image_loss_meter.avg, text_loss_meter.avg))
 
-        print('====> Epoch: {} Joint loss: {:.4f}\tImage loss: {:.4f}\tText loss: {:.4f}'.format(
+        print('====> Epoch: {}\tJoint loss: {:.4f}\tImage loss: {:.4f}\tText loss: {:.4f}'.format(
             epoch, joint_loss_meter.avg, image_loss_meter.avg, text_loss_meter.avg))
 
 
@@ -203,7 +203,7 @@ if __name__ == "__main__":
         test_text_loss /= len(test_loader.dataset)
         test_loss /= len(test_loader.dataset)
         
-        print('====> Test joint loss: {:.4f}\timage loss: {:.4f}\ttext loss:{:.4f}'.format(
+        print('====> Test Epoch\tJoint loss: {:.4f}\tImage loss: {:.4f}\tText loss:{:.4f}'.format(
             test_joint_loss, test_image_loss, test_text_loss))
         
         return test_loss, (test_joint_loss, test_image_loss, test_text_loss)

### 2017-10-22 mnist/weak.py
Bugfixes to new weak.py design

@@ -87,7 +87,9 @@ def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, 
             loss_3 = text_loss_function(recon_text_3, text, mu_3, logvar_3,
                                         batch_size=batch_size)  
             loss = loss_2 + loss_3          
-
+            image_loss_meter.update(loss_2.data[0], len(image))
+            text_loss_meter.update(loss_3.data[0], len(text))
+            
             # if we flip(weak_perc), then we show a paired relation example.
             flip = np.random.random()
             if flip < weak_perc:
@@ -95,11 +97,9 @@ def train_pipeline(out_dir, weak_perc, n_latents=20, batch_size=128, epochs=20, 
                 loss_1 = joint_loss_function(recon_image_1, image, recon_text_1, text, mu_1, logvar_1,
                                              batch_size=args.batch_size)
                 loss = loss + loss_1
+                joint_loss_meter.update(loss_1.data[0], len(image))
 
             loss.backward()
-            joint_loss_meter.update(loss_1.data[0], len(image))
-            image_loss_meter.update(loss_2.data[0], len(image))
-            text_loss_meter.update(loss_3.data[0], len(text))
             optimizer.step()
 
             if batch_idx % log_interval == 0:

### 2017-10-22 multimnist/dataset.py
Added encoder and decoder networks

@@ -1,106 +1,0 @@
-"""
-This script generates a dataset similar to the Multi-MNIST dataset
-described in [1].
-
-[1] Eslami, SM Ali, et al. "Attend, infer, repeat: Fast scene
-understanding with generative models." Advances in Neural Information
-Processing Systems. 2016.
-"""
-
-from __future__ import division
-from __future__ import print_function
-from __future__ import absolute_import
-
-import os
-import numpy as np
-import torch
-import torchvision.datasets as dset
-from scipy.misc import imresize
-
-
-def sample_one(canvas_size, mnist):
-    i = np.random.randint(mnist['digits'].shape[0])
-    digit = mnist['digits'][i]
-    label = mnist['labels'][i]
-    scale = 0.1 * np.random.randn() + 1.3
-    resized = imresize(digit, 1. / scale)
-    w = resized.shape[0]
-    assert w == resized.shape[1]
-    padding = canvas_size - w
-    pad_l = np.random.randint(0, padding)
-    pad_r = np.random.randint(0, padding)
-    pad_width = ((pad_l, padding - pad_l), (pad_r, padding - pad_r))
-    positioned = np.pad(resized, pad_width, 'constant', constant_values=0)
-    return positioned, label
-
-
-def sample_multi(num_digits, canvas_size, mnist):
-    canvas = np.zeros((canvas_size, canvas_size))
-    labels = []
-    for _ in range(num_digits):
-        positioned_digit, label = sample_one(canvas_size, mnist)
-        canvas += positioned_digit
-        labels.append(label)
-    
-    # Crude check for overlapping digits.
-    if np.max(canvas) > 255:
-        return sample_multi(num_digits, canvas_size, mnist)
-    else:
-        return canvas, labels
-
-
-def mk_dataset(n, mnist, min_digits, max_digits, canvas_size):
-    x = []
-    y = []
-    for _ in range(n):
-        num_digits = np.random.randint(min_digits, max_digits + 1)
-        canvas, labels = sample_multi(num_digits, canvas_size, mnist)
-        x.append(canvas)
-        y.append(labels)
-    return np.array(x, dtype=np.uint8), y
-
-
-def load_mnist():
-    train_loader = torch.utils.data.DataLoader(
-        dset.MNIST(root='./data', train=True, download=True))
-
-    test_loader = torch.utils.data.DataLoader(
-        dset.MNIST(root='./data', train=False, download=True))
-    
-    train_data = {
-        'digits': train_loader.dataset.train_data.numpy(),
[…]

### 2017-10-22 multimnist/datasets.py

@@ -1,0 +1,177 @@
+"""
+This script generates a dataset similar to the Multi-MNIST dataset
+described in [1].
+
+[1] Eslami, SM Ali, et al. "Attend, infer, repeat: Fast scene
+understanding with generative models." Advances in Neural Information
+Processing Systems. 2016.
+"""
+
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import os
+import numpy as np
+from scipy.misc import imresize
+
+import torch
+import torchvision.datasets as dset
+from torch.utils.data.dataset import Dataset
+
+
+class MultiMNIST(Dataset):
+    """images with 0 to 2 images of non-overlapping MNIST numbers."""
+    processed_folder = 'multimnist'
+    training_file = 'training.pt'
+    test_file = 'test.pt'
+
+    def __init__(self, root, train=True, transform=None, target_transform=None, download=False):
+        self.root = os.path.expanduser(root)
+        self.transform = transform
+        self.target_transform = target_transform
+        self.train = train  # training set or test set
+
+        if download:
+            self.download()
+
+        if not self._check_exists():
+            raise RuntimeError('Dataset not found.' +
+                               ' You can use download=True to download it')
+
+        if self.train:
+            self.train_data, self.train_labels = torch.load(
+                os.path.join(self.root, self.processed_folder, self.training_file))
+        else:
+            self.test_data, self.test_labels = torch.load(
+                os.path.join(self.root, self.processed_folder, self.test_file))
+
+    def __getitem__(self, index):
+        """
+        Args:
+            index (int): Index
+        Returns:
+            tuple: (image, target) where target is index of the target class.
+        """
+        if self.train:
+            img, target = self.train_data[index], self.train_labels[index]
+        else:
+            img, target = self.test_data[index], self.test_labels[index]
+
+        # doing this so that it is consistent with all other datasets
+        # to return a PIL Image
+        img = Image.fromarray(img.numpy(), mode='L')
+
+        if self.transform is not None:
+            img = self.transform(img)
+
+        if self.target_transform is not None:
+            target = self.target_transform(target)
+
+        return img, target
+
+    def __len__(self):
+        if self.train:
[…]

### 2017-10-22 multimnist/model.py

@@ -1,0 +1,195 @@
+"""This model will be quite similar to mnist/model.py 
+except we will need to be slightly fancier in the 
+encoder/decoders for each modality. Likely, we will need 
+convolutions/deconvolutions and RNNs.
+"""
+
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import torch
+import torch.nn as nn
+from torch.autograd import Variable
+from torch.nn import functional as F
+from torch.nn.parameter import Parameter
+
+from utils import n_characters
+from utils import SOS, EOS
+
+
+class ImageEncoder(nn.Module):
+    """This task is quite a bit harder than MNIST so we probably need 
+    to use an RNN of some form. This will be good to get us ready for
+    natural images.
+
+    :param n_latents: size of latent vector
+    """
+    def __init__(self, n_latents):
+        super(ImageEncoder, self).__init__()
+        self.features = nn.Sequential(
+            nn.Conv2d(1, 10, kernel_size=5),
+            nn.MaxPool2d(2),
+            nn.ReLU(),
+            nn.BatchNorm2d(10),
+            nn.Conv2d(10, 20, kernel_size=5),
+            nn.MaxPool2d(2),
+            nn.ReLU(),
+            nn.BatchNorm2d(20),
+        )
+        self.classifier = nn.Sequential(
+            nn.Linear(320, 50),
+            nn.ReLU(),
+            nn.Dropout(p=0.1),
+            nn.Linear(50, n_latents * 2)
+        )
+        self.n_latents = n_latents
+
+    def forward(self, x)
+        n_latents = self.n_latents
+        x = self.features(x)
+        x = x.view(-1)
+        x = self.classifier(x)
+        return x[:, :n_latents], x[:, n_latents:]
+
+
+class TextEncoder(nn.Module):
+    """Given variable length text, we move past a single embedding 
+    to a bidirectional RNN.
+
+    :param n_latents: size of latent vector
+    :param n_characters: number of possible characters (10 for MNIST)
+    """
+    def __init__(self, n_latents, n_characters):
+        super(TextEncoder, self).__init__()
+        self.embed = nn.Embedding(n_characters, 50)
+        self.gru = nn.GRU(50, 50, 1, dropout=0.1, bidirectional=True)
+        self.h2p = nn.Linear(50, n_latents * 2)  # hiddens to parameters
[…]

### 2017-10-22 multimnist/model.py
Added scripts to train imageonly and textonly in multimnist case

@@ -18,6 +18,127 @@ from utils import n_characters
 from utils import SOS, EOS
 
 
+class MultimodalVAE(nn.Module):
+    def __init__(self, n_latents=20):
+        super(MultimodalVAE, self).__init__()
+        self.image_encoder = ImageEncoder(n_latents)
+        self.image_decoder = ImageDecoder(n_latents)
+        self.text_encoder = TextEncoder(n_latents)
+        self.text_decoder = TextDecoder(n_latents)
+        self.experts = ProductOfExperts()
+
+    def reparametrize(self, mu, logvar):
+        if self.training:
+            std = logvar.mul(0.5).exp_()
+            eps = Variable(std.data.new(std.size()).normal_())
+            return eps.mul(std).add_(mu)
+        else:  # return mean during inference
+            return mu
+
+    def encode_image(self, x):
+        return self.image_encoder(x)
+
+    def decode_image(self, x):
+        return self.image_decoder(x)
+
+    def encode_text(self, x):
+        return self.text_encoder(x)
+
+    def decode_text(self, x):
+        return self.text_decoder(x)
+
+    def forward(self, image=None, text=None):
+        # can't just put nothing
+        assert image is not None or text is not None
+        
+        if image is not None and text is not None:
+            # compute separate gaussians per modality
+            image_mu, image_logvar = self.encode_image(image)
+            text_mu, text_logvar = self.encode_text(text)
+            mu = torch.stack((image_mu, text_mu), dim=0)
+            logvar = torch.stack((image_logvar, text_logvar), dim=0)
+            # grab learned mixture weights and sample
+            mu, logvar = self.experts(mu, logvar)
+        elif image is not None:
+            mu, logvar = self.encode_image(image)
+        elif text is not None:
+            mu, logvar = self.encode_text(text)
+        
+        # reparametrization trick to sample
+        z = self.reparametrize(mu, logvar)
+        # reconstruct inputs based on that gaussian
+        image_recon = self.decode_image(z)
+        text_recon = self.decode_text(z)
+
+        return image_recon, text_recon, mu, logvar
+
+    def gen_latents(self, image, text):
+        # compute separate gaussians per modality
+        image_mu, image_logvar = self.encode_image(image)
+        text_mu, text_logvar = self.encode_text(text)
+        mu = torch.stack((image_mu, text_mu), dim=0)
[…]

### 2017-10-22 multimnist/train_textonly.py

@@ -1,0 +1,157 @@
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import os
+import shutil
+import sys
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.autograd import Variable
+from torchvision import transforms
+
+import datasets
+from model import TextVAE
+from train import AverageMeter
+
+
+def save_checkpoint(state, is_best, folder='./', filename='checkpoint.pth.tar'):
+    torch.save(state, os.path.join(folder, filename))
+    if is_best:
+        shutil.copyfile(os.path.join(folder, filename),
+                        os.path.join(folder, 'model_best.pth.tar'))
+
+
+def load_checkpoint(file_path, use_cuda=False):
+    """Return EmbedNet instance"""
+    if use_cuda:
+        checkpoint = torch.load(file_path)
+    else:
+        checkpoint = torch.load(file_path,
+                                map_location=lambda storage, location: storage)
+
+    vae = TextVAE(n_latents=checkpoint['n_latents'])
+    vae.load_state_dict(checkpoint['state_dict'])
+    
+    return vae
+
+
+def loss_function(recon_x, x, mu, logvar, batch_size=128):
+    BCE = F.nll_loss(recon_text, text)
+    # see Appendix B from VAE paper:
+    # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
+    # https://arxiv.org/abs/1312.6114
+    # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
+    KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
+    KLD /= batch_size * 50  # for each embedding unit
+    return BCE + KLD
+
+
+if __name__ == "__main__":
+    import argparse
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--n_latents', type=int, default=20,
+                        help='size of the latent embedding')
+    parser.add_argument('--batch_size', type=int, default=128, metavar='N',
+                        help='input batch size for training (default: 128)')
+    parser.add_argument('--epochs', type=int, default=20, metavar='N',
+                        help='number of epochs to train (default: 20)')
+    parser.add_argument('--lr', type=float, default=1e-3, metavar='LR',
+                        help='learning rate (default: 1e-3)')
+    parser.add_argument('--log_interval', type=int, default=10, metavar='N',
[…]

### 2017-10-22 mnist/test_weak.py
fixed test_weak plot function

@@ -2,6 +2,8 @@ from __future__ import division
 from __future__ import print_function
 from __future__ import absolute_import
 
+import numpy as np
+
 import torch
 from torchvision import datasets, transforms
 
@@ -41,6 +43,9 @@ if __name__ == "__main__":
         y.append(weak_acc)
         print('Got accuracies for %s.' % dir_path)
 
+    x, y = np.array(x), np.array(y)
+    ix = np.argsort(x)
+    x, y = x[ix], y[ix]
 
     plt.figure()
     plt.plot(x, y, '-o')

### 2017-10-22 multimnist/datasets.py
bugfixes in datasets.py and model.py

@@ -158,8 +158,8 @@ def make_dataset(root, folder, training_file, test_file):
     train_x, train_y = mk_dataset(60000, train_mnist, 0, 2, 50)
     test_x, test_y = mk_dataset(10000, test_mnist, 0, 2, 50)
     
-    train_x, train_y = torch.load(train_x).byte(), torch.load(train_y).long()
-    test_x, test_y = torch.load(test_x).byte(), torch.load(test_y).long()
+    train_x = torch.from_numpy(train_x).byte()
+    test_x = torch.from_numpy(test_x).byte()
 
     training_set = (train_x, train_y)
     test_set = (test_x, test_y)

### 2017-10-22 multimnist/model.py

@@ -166,7 +166,7 @@ class ImageEncoder(nn.Module):
         )
         self.n_latents = n_latents
 
-    def forward(self, x)
+    def forward(self, x):
         n_latents = self.n_latents
         x = self.features(x)
         x = x.view(-1)

### 2017-11-16 mnist/manifold.py
Update manifold (MNIST)

@@ -59,9 +59,11 @@ if __name__ == "__main__":
     with open('./results/dump.pkl', 'wb') as fp:
         pickle.dump({'latents': latents, 'labels': labels}, fp)
 
-    if latents.shape[1] > 2:
-        pca = PCA(n_components=2)
-        latents = pca.fit_transform(latents)
+    # perform staggered dimensionality reduction
+    pca_50 = PCA(n_components=50)
+    latents = pca_50.fit_transform(latents)
+    tsne = TSNE(n_components=2, verbose=1, perplexity=40, n_iter=300)
+    latents = tsne.fit_transform(latents)    
 
     # now we have latents guaranteed to be 2 dimensions
     # let's plot the manifold

### 2017-11-16 mnist/loglikelihood.py
Added script to compute loglikelihood of image only

@@ -1,4 +1,4 @@
-"""Compute log-likelihoods for entire dataset"""
+"""Compute log-likelihoods for test dataset"""
 
 from __future__ import division
 from __future__ import print_function

### 2017-11-16 mnist/loglikelihood_imageonly.py

@@ -1,0 +1,75 @@
+"""Compute log-likelihoods for test dataset 
+but for image_only VAE.
+"""
+
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import torch
+import torch.nn.functional as F
+from torch.autograd import Variable
+from torchvision import datasets, transforms
+
+from train import load_checkpoint
+
+
+def compute_nll(model, loader, n_samples=1, use_cuda=False):
+    model.eval()
+    test_nll = 0
+
+    for batch_idx, (image, _) in enumerate(loader):
+        if use_cuda:
+            image = image.cuda()
+        image = Variable(image, volatile=True)
+        image = image.view(-1, 784)
+
+        _, mu, logvar = model(image)
+        batch_size, n_latents = mu.size(0), mu.size(1)
+        sample = Variable(torch.randn(n_samples, n_latents))
+
+        if use_cuda:
+            sample = sample.cuda()
+
+        std = logvar.mul(0.5).exp_()
+        mu = mu.unsqueeze(1).repeat(1, n_samples, 1)
+        std = std.unsqueeze(1).repeat(1, n_samples, 1)
+
+        sample = sample.unsqueeze(0).repeat(batch_size, 1, 1)
+        sample = sample.mul(std).add_(mu)
+
+        nll = 0
+        for i in xrange(n_samples):
+            recon_image = model.decode_image(sample[:, i])
+            nll += F.binary_cross_entropy(recon_image, image, size_average=False)
+
+        test_nll += (nll / n_samples)
+        print('Evaluating: [{}/{} ({:.0f}%)]'.format(batch_idx * len(image), len(loader.dataset),
+                                                     100. * batch_idx / len(loader)))   
+
+    test_nll /= len(loader.dataset)
+    return test_nll
+
+
+if __name__ == "__main__":
+    import argparse
+    parser = argparse.ArgumentParser()
+    parser.add_argument('model_path', type=str, help='path to trained model file')
+    parser.add_argument('--n_samples', type=int, default=100, 
+                        help='number of samples to use to estimate the ELBO')
+    parser.add_argument('--cuda', action='store_true', default=False,
+                        help='enables CUDA training')
+    args = parser.parse_args()
+    args.cuda = args.cuda and torch.cuda.is_available()
+
+    # loader for MNIST
+    loader = torch.utils.data.DataLoader(
[…]

### 2017-11-16 mnist/manifold.py

@@ -59,11 +59,13 @@ if __name__ == "__main__":
     with open('./results/dump.pkl', 'wb') as fp:
         pickle.dump({'latents': latents, 'labels': labels}, fp)
 
-    # perform staggered dimensionality reduction
-    pca_50 = PCA(n_components=50)
-    latents = pca_50.fit_transform(latents)
+    # > 50 dimensions is too expensive for tSNE
+    if latents.size(1) > 50:  
+        pca_50 = PCA(n_components=50)
+        latents = pca_50.fit_transform(latents)
+    
     tsne = TSNE(n_components=2, verbose=1, perplexity=40, n_iter=300)
-    latents = tsne.fit_transform(latents)    
+    latents = tsne.fit_transform(latents)
 
     # now we have latents guaranteed to be 2 dimensions
     # let's plot the manifold

### 2017-11-16 mnist/loglikelihood_imageonly.py
Bugfixes in MNIST scripts

@@ -11,7 +11,7 @@ import torch.nn.functional as F
 from torch.autograd import Variable
 from torchvision import datasets, transforms
 
-from train import load_checkpoint
+from train_imageonly import load_checkpoint
 
 
 def compute_nll(model, loader, n_samples=1, use_cuda=False):
@@ -40,7 +40,7 @@ def compute_nll(model, loader, n_samples=1, use_cuda=False):
 
         nll = 0
         for i in xrange(n_samples):
-            recon_image = model.decode_image(sample[:, i])
+            recon_image = model.decode(sample[:, i])
             nll += F.binary_cross_entropy(recon_image, image, size_average=False)
 
         test_nll += (nll / n_samples)
@@ -70,6 +70,8 @@ if __name__ == "__main__":
 
     vae = load_checkpoint(args.model_path, use_cuda=args.cuda)
     vae.eval()
+    if args.cuda:
+        vae.cuda()
 
     nll = compute_nll(vae, loader, use_cuda=args.cuda, n_samples=args.n_samples)
-    print('\nTest NLL: {:.4f}'.format(nll))
+    print('\nTest NLL: {:.4f}'.format(nll.data[0]))

### 2017-11-16 mnist/sample.py

@@ -45,7 +45,6 @@ if __name__ == "__main__":
     import argparse
     parser = argparse.ArgumentParser()
     parser.add_argument('n_samples', type=int, help='Number of images and texts to sample.')
-    parser.add_argument('--n_latents', type=int, default=20, help='Dimension of latent embedding.')
     parser.add_argument('--condition_on_image', type=int, default=None,
                         help='If True, generate text conditioned on an image.')
     parser.add_argument('--condition_on_text', type=int, default=None, 
@@ -56,8 +55,7 @@ if __name__ == "__main__":
     args.cuda = args.cuda and torch.cuda.is_available()
 
     # load trained model
-    vae = load_checkpoint('./trained_models/model_best.pth.tar', 
-                          n_latents=args.n_latents, use_cuda=args.cuda)
+    vae = load_checkpoint('./trained_models/model_best.pth.tar', use_cuda=args.cuda)
     vae.eval()
     if args.cuda:
         vae.cuda()
@@ -98,7 +96,8 @@ if __name__ == "__main__":
         std = logvar.mul(0.5).exp_()
 
     # sample from uniform gaussian
-    sample = Variable(torch.randn(args.n_samples, args.n_latents))
+    n_latents = vae.encoder.n_latents
+    sample = Variable(torch.randn(args.n_samples, n_latents))
     if args.cuda:
         sample = sample.cuda()
     

### 2017-11-16 mnist/sample_imageonly.py
Add scripts to sample MNIST images from

@@ -1,0 +1,92 @@
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import numpy as np
+
+import torch
+from torch.autograd import Variable
+from torchvision import datasets, transforms
+from torchvision.utils import save_image
+
+from train_imageonly import load_checkpoint
+
+
+def fetch_mnist_image(_label):
+    # find an example of the image in our dataset
+    loader = torch.utils.data.DataLoader(
+        datasets.MNIST('./data', train=False, download=True, 
+                       transform=transforms.ToTensor()),
+        batch_size=128, shuffle=False)
+    # load all data into a single structure
+    images, labels = [], []
+    for batch_idx, (image, label) in enumerate(loader):
+        image = image.view(-1, 784)
+        images.append(image)
+        labels.append(label)
+    images = torch.cat(images).cpu().numpy()
+    labels = torch.cat(labels).cpu().numpy()
+    # take all the ones where it's an image of the correct label
+    images = images[labels == _label]
+    # randomly choose one
+    image = images[np.random.choice(np.arange(images.shape[0]))]
+    image = torch.from_numpy(image).float() 
+    image = image.unsqueeze(0)
+    return Variable(image, volatile=True)
+
+
+if __name__ == "__main__":
+    import os
+    import argparse
+    parser = argparse.ArgumentParser()
+    parser.add_argument('n_samples', type=int, help='Number of images and texts to sample.')
+    parser.add_argument('--condition_on_image', type=int, default=None,
+                        help='If True, generate text conditioned on an image.')
+    parser.add_argument('--cuda', action='store_true', default=False,
+                        help='enables CUDA training')
+    args = parser.parse_args()
+    args.cuda = args.cuda and torch.cuda.is_available()
+
+    # load trained model
+    vae = load_checkpoint('./trained_models/model_best.pth.tar', use_cuda=args.cuda)
+    vae.eval()
+    if args.cuda:
+        vae.cuda()
+
+    # mode 1: unconditional generation
+    if not args.condition_on_image:
+        mu = Variable(torch.Tensor([0]))
+        std = Variable(torch.Tensor([1]))
+    else:  # mode 2: generate conditioned on image
+        image = fetch_mnist_image(args.condition_on_image)
+        if args.cuda:
+            image = image.cuda()
+        mu, logvar = vae.encode_image(image)
[…]

### 2017-11-16 mnist/train.py

@@ -41,14 +41,14 @@ def save_checkpoint(state, is_best, folder='./', filename='checkpoint.pth.tar'):
                         os.path.join(folder, 'model_best.pth.tar'))
 
 
-def load_checkpoint(file_path, n_latents=20, use_cuda=False):
+def load_checkpoint(file_path, use_cuda=False):
     """Return EmbedNet instance"""
     if use_cuda:
         checkpoint = torch.load(file_path)
     else:
         checkpoint = torch.load(file_path,
                                 map_location=lambda storage, location: storage)
-
+    n_latents = checkpoint['n_latents']
     vae = MultimodalVAE(n_latents=n_latents)
     vae.load_state_dict(checkpoint['state_dict'])
     

### 2017-11-16 mnist/train_imageonly.py

@@ -49,8 +49,8 @@ def load_checkpoint(file_path, use_cuda=False):
     else:
         checkpoint = torch.load(file_path,
                                 map_location=lambda storage, location: storage)
-
-    vae = VAE()
+    n_latents = checkpoint['n_latents']
+    vae = VAE(n_latents=n_latents)
     vae.load_state_dict(checkpoint['state_dict'])
     
     return vae
@@ -152,6 +152,7 @@ if __name__ == "__main__":
         save_checkpoint({
             'state_dict': vae.state_dict(),
             'best_loss': best_loss,
+            'n_latents': args.n_latents,
             'optimizer' : optimizer.state_dict(),
         }, is_best, folder='./trained_models/image_only')
 

### 2017-11-17 mnist/imageonly/loglikelihood_imageonly.py
Move image_only files to its own folder

@@ -64,7 +64,7 @@ if __name__ == "__main__":
 
     # loader for MNIST
     loader = torch.utils.data.DataLoader(
-        datasets.MNIST('./data', train=False, download=True,
+        datasets.MNIST('../data', train=False, download=True,
                        transform=transforms.ToTensor()),
         batch_size=64, shuffle=True)
 

### 2017-11-17 mnist/imageonly/manifold_imageonly.py

@@ -6,8 +6,9 @@ import torch
 from torch.autograd import Variable
 from torchvision import transforms, datasets
 
-from train import load_checkpoint
+from train_imageonly import load_checkpoint
 from sklearn.decomposition import PCA
+from sklearn.manifold import TSNE
 
 
 if __name__ == "__main__":
@@ -30,7 +31,7 @@ if __name__ == "__main__":
     # use this instead of shuffle loader because we don't want 
     # to have negative examples here.
     loader = torch.utils.data.DataLoader(
-        datasets.MNIST('./data', train=False, download=True, 
+        datasets.MNIST('../data', train=False, download=True, 
                        transform=transforms.ToTensor()),
         batch_size=128, shuffle=True)
 
@@ -45,7 +46,9 @@ if __name__ == "__main__":
         image, text = Variable(image, volatile=True), Variable(text, volatile=True)
         image = image.view(-1, 784)  # flatten image
 
-        z = vae.gen_latents(image, text)
+        mu, logvar = vae.encode(image)
+        z = vae.reparameterize(mu, logvar)
+        
         if batch_idx == 0:
             latents = z
             labels = text
@@ -60,7 +63,7 @@ if __name__ == "__main__":
         pickle.dump({'latents': latents, 'labels': labels}, fp)
 
     # > 50 dimensions is too expensive for tSNE
-    if latents.size(1) > 50:  
+    if latents.shape[1] > 50:  
         pca_50 = PCA(n_components=50)
         latents = pca_50.fit_transform(latents)
     

### 2017-11-17 mnist/imageonly/sample_imageonly.py

@@ -39,7 +39,9 @@ if __name__ == "__main__":
     import os
     import argparse
     parser = argparse.ArgumentParser()
-    parser.add_argument('n_samples', type=int, help='Number of images and texts to sample.')
+    parser.add_argument('model_path', type=str, help='path to model file')
+    parser.add_argument('--n_samples', type=int, default=64, 
+                        help='Number of images and texts to sample.')
     parser.add_argument('--condition_on_image', type=int, default=None,
                         help='If True, generate text conditioned on an image.')
     parser.add_argument('--cuda', action='store_true', default=False,
@@ -48,7 +50,7 @@ if __name__ == "__main__":
     args.cuda = args.cuda and torch.cuda.is_available()
 
     # load trained model
-    vae = load_checkpoint('./trained_models/model_best.pth.tar', use_cuda=args.cuda)
+    vae = load_checkpoint(args.model_path, use_cuda=args.cuda)
     vae.eval()
     if args.cuda:
         vae.cuda()
@@ -57,6 +59,9 @@ if __name__ == "__main__":
     if not args.condition_on_image:
         mu = Variable(torch.Tensor([0]))
         std = Variable(torch.Tensor([1]))
+        if args.cuda:
+            mu = mu.cuda()
+            std = std.cuda()
     else:  # mode 2: generate conditioned on image
         image = fetch_mnist_image(args.condition_on_image)
         if args.cuda:
@@ -65,7 +70,7 @@ if __name__ == "__main__":
         std = logvar.mul(0.5).exp_()
 
     # sample from uniform gaussian
-    n_latents = vae.encoder.n_latents
+    n_latents = vae.n_latents
     sample = Variable(torch.randn(args.n_samples, n_latents))
     if args.cuda:
         sample = sample.cuda()
@@ -84,9 +89,3 @@ if __name__ == "__main__":
     save_image(image_recon.data.view(args.n_samples, 1, 28, 28),
                './results/sample_image.png')
 
-    # save text samples to filesystem
-    with open('./results/sample_texts.txt', 'w') as fp:
-        text_recon_np = text_recon.data.numpy()
-        text_recon_np = np.argmax(text_recon_np, axis=1).tolist()
-        for i, item in enumerate(text_recon_np):
-            fp.write('Text (%d): %s\n' % (i, item))

### 2017-11-17 mnist/imageonly/train_imageonly.py

@@ -14,6 +14,7 @@ from torch.autograd import Variable
 from torchvision import transforms, datasets
 from torchvision.utils import save_image
 
+sys.path.append('..')
 from model import VAE
 
 
@@ -88,12 +89,18 @@ if __name__ == "__main__":
     args = parser.parse_args()
     args.cuda = args.cuda and torch.cuda.is_available()
 
+    if not os.path.isdir('./trained_models'):
+        os.mkdir('./trained_models')
+
+    if not os.path.isdir('./results'):
+        os.mkdir('./results')
+
     train_loader = torch.utils.data.DataLoader(
-        datasets.MNIST('./data', train=True, download=True,
+        datasets.MNIST('../data', train=True, download=True,
                        transform=transforms.ToTensor()),
         batch_size=args.batch_size, shuffle=True)
     test_loader = torch.utils.data.DataLoader(
-        datasets.MNIST('./data', train=False, download=True, 
+        datasets.MNIST('../data', train=False, download=True, 
                        transform=transforms.ToTensor()),
         batch_size=args.batch_size, shuffle=True)
 
@@ -154,7 +161,7 @@ if __name__ == "__main__":
             'best_loss': best_loss,
             'n_latents': args.n_latents,
             'optimizer' : optimizer.state_dict(),
-        }, is_best, folder='./trained_models/image_only')
+        }, is_best, folder='./trained_models')
 
         if is_best:
             sample = Variable(torch.randn(64, 20))
@@ -163,4 +170,4 @@ if __name__ == "__main__":
 
             sample = vae.decode(sample).cpu()
             save_image(sample.data.view(64, 1, 28, 28),
-                       './results/image_only/sample_image.png')
+                       './results/sample_image.png')

### 2017-11-17 mnist/model.py

@@ -17,6 +17,7 @@ class MultimodalVAE(nn.Module):
         self.text_encoder = TextEncoder(n_latents)
         self.text_decoder = TextDecoder(n_latents)
         self.experts = ProductOfExperts()
+        self.n_latents = n_latents
 
     def reparametrize(self, mu, logvar):
         if self.training:

### 2017-11-17 mnist/sample.py

@@ -44,7 +44,9 @@ if __name__ == "__main__":
     import os
     import argparse
     parser = argparse.ArgumentParser()
-    parser.add_argument('n_samples', type=int, help='Number of images and texts to sample.')
+    parser.add_argument('--model_path', type=str, help='path to trained model file')
+    parser.add_argument('--n_samples', type=int, default=64, 
+                        help='Number of images and texts to sample.')
     parser.add_argument('--condition_on_image', type=int, default=None,
                         help='If True, generate text conditioned on an image.')
     parser.add_argument('--condition_on_text', type=int, default=None, 

### 2017-11-17 mnist/imageonly/manifold_imageonly.py
Added script for PCA only cuz its faster

@@ -23,6 +23,7 @@ if __name__ == "__main__":
 
     parser = argparse.ArgumentParser()
     parser.add_argument('model_path', type=str, help='path to trained model file')
+    parser.add_argument('--pca_only', action='store_true', default=False)
     parser.add_argument('--cuda', action='store_true', default=False,
                         help='enables CUDA training')
     args = parser.parse_args()
@@ -62,13 +63,16 @@ if __name__ == "__main__":
     with open('./results/dump.pkl', 'wb') as fp:
         pickle.dump({'latents': latents, 'labels': labels}, fp)
 
-    # > 50 dimensions is too expensive for tSNE
-    if latents.shape[1] > 50:  
-        pca_50 = PCA(n_components=50)
-        latents = pca_50.fit_transform(latents)
-    
-    tsne = TSNE(n_components=2, verbose=1, perplexity=40, n_iter=300)
-    latents = tsne.fit_transform(latents)
+    if args.pca_only:
+        pca_2 = PCA(n_components=2)
+        latents = pca_2.fit_transform(latents)
+    else:
+        # > 50 dimensions is too expensive for tSNE
+        if latents.shape[1] > 50:  
+            pca_50 = PCA(n_components=50)
+            latents = pca_50.fit_transform(latents)
+        tsne = TSNE(n_components=2, verbose=1, perplexity=40, n_iter=300)
+        latents = tsne.fit_transform(latents)
 
     # now we have latents guaranteed to be 2 dimensions
     # let's plot the manifold

### 2017-11-17 coco/train.py
Lots of exploring with hyperparameters leads to better understanding of what size_average does and how to weight between likelihood terms

@@ -70,18 +70,18 @@ def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, tex
     
     if recon_image is not None and image is not None:
         image_BCE = lambda_xy * F.binary_cross_entropy(recon_image.view(-1, 1 * 64 * 64), 
-                                                       image.view(-1, 1 * 64 * 64), size_average=False)
+                                                       image.view(-1, 1 * 64 * 64))
 
     if recon_text is not None and text is not None:
-        text_BCE = lambda_yx * F.nll_loss(recon_text.view(-1, recon_text.size(2)), text.view(-1),
-                                          size_average=False)
+        text_BCE = lambda_yx * F.nll_loss(recon_text.view(-1, recon_text.size(2)), text.view(-1))
 
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
     # https://arxiv.org/abs/1312.6114
     # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-    return (image_BCE + text_BCE + KLD * kl_lambda) / batch_size
+    KLD = KLD / batch_size * kl_lambda
+    return image_BCE + text_BCE + KLD
 
 
 if __name__ == "__main__":

### 2017-11-17 mnist/imageonly/latent_viz_imageonly.py

@@ -14,7 +14,7 @@ import torch
 from torch.autograd import Variable
 
 import numpy as np
-from train import load_checkpoint
+from train_imageonly import load_checkpoint
 
 
 def array2d_string(A):
@@ -42,7 +42,7 @@ if __name__ == "__main__":
 
     vae = load_checkpoint(args.model_path, use_cuda=args.cuda)
     vae.eval()
-
+    
     if args.cuda:
         vae.cuda()
 
@@ -51,7 +51,6 @@ if __name__ == "__main__":
     y_values = torch.linspace(-3, 3, ny)
 
     image_canvas = torch.zeros((28 * ny, 28 * nx))
-    text_canvas = torch.zeros((ny, nx))
 
     for i, yi in enumerate(x_values):
         for j, xi in enumerate(y_values):
@@ -59,12 +58,9 @@ if __name__ == "__main__":
             if args.cuda:
                 data = data.cuda()
 
-            recon_image = vae.decode_image(data)
-            recon_text = vae.decode_text(data)
-
+            recon_image = vae.decode(data)
             image_canvas[(nx-i-1)*28:(nx-i)*28, j*28:(j+1)*28] = \
                 recon_image[0].cpu().data.reshape(28, 28)
-            text_canvas[i, j] = recon_text[0].cpu().data
 
     if not os.path.exists('./results'):
         os.makedirs('./results')
@@ -75,6 +71,3 @@ if __name__ == "__main__":
     plt.tight_layout()
     plt.savefig('./results/latent_viz_image.png')
 
-    text_canvas = text_canvas.numpy().tolist()
-    with open('./results/latent_viz_text.txt', 'wb') as fp:
-        fp.write(array2d_string(text_canvas))

### 2017-11-17 mnist/imageonly/train_imageonly.py

@@ -59,7 +59,7 @@ def load_checkpoint(file_path, use_cuda=False):
 
 def loss_function(recon_x, x, mu, logvar):
     batch_size = recon_x.size(0)
-    BCE = F.binary_cross_entropy(recon_x, x.view(-1, 784), size_average=False)
+    BCE = F.binary_cross_entropy(recon_x, x.view(-1, 784))
 
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014

### 2017-11-17 mnist/manifold.py

@@ -22,6 +22,7 @@ if __name__ == "__main__":
 
     parser = argparse.ArgumentParser()
     parser.add_argument('model_path', type=str, help='path to trained model file')
+    parser.add_argument('--pca_only', action='store_true', default=False)
     parser.add_argument('--cuda', action='store_true', default=False,
                         help='enables CUDA training')
     args = parser.parse_args()
@@ -59,13 +60,16 @@ if __name__ == "__main__":
     with open('./results/dump.pkl', 'wb') as fp:
         pickle.dump({'latents': latents, 'labels': labels}, fp)
 
-    # > 50 dimensions is too expensive for tSNE
-    if latents.size(1) > 50:  
-        pca_50 = PCA(n_components=50)
-        latents = pca_50.fit_transform(latents)
+    if args.pca_only:
+        pca_2 = PCA(n_components=2)
+        latents = pca_2.fit_transform(latents)
+    else:
+        if latents.size(1) > 50:  
+            pca_50 = PCA(n_components=50)
+            latents = pca_50.fit_transform(latents)
     
-    tsne = TSNE(n_components=2, verbose=1, perplexity=40, n_iter=300)
-    latents = tsne.fit_transform(latents)
+        tsne = TSNE(n_components=2, verbose=1, perplexity=40, n_iter=300)
+        latents = tsne.fit_transform(latents)
 
     # now we have latents guaranteed to be 2 dimensions
     # let's plot the manifold

### 2017-11-17 mnist/model.py

@@ -39,6 +39,15 @@ class MultimodalVAE(nn.Module):
     def decode_text(self, x):
         return self.text_decoder(x)
 
+    def prior(self, size, use_cuda=False):
+        mu = Variable(torch.zeros(size))
+        logvar = Variable(torch.log(torch.ones(size)))
+        if use_cuda:
+            mu = mu.cuda()
+            logvar = logvar.cuda()
+
+        return mu, logvar
+
     def forward(self, image=None, text=None):
         # can't just put nothing
         assert image is not None or text is not None
@@ -49,13 +58,21 @@ class MultimodalVAE(nn.Module):
             text_mu, text_logvar = self.encode_text(text)
             mu = torch.stack((image_mu, text_mu), dim=0)
             logvar = torch.stack((image_logvar, text_logvar), dim=0)
-            # grab learned mixture weights and sample
-            mu, logvar = self.experts(mu, logvar)
         elif image is not None:
             mu, logvar = self.encode_image(image)
+            mu, logvar = mu.unsqueeze(0), logvar.unsqueeze(0)
         elif text is not None:
             mu, logvar = self.encode_text(text)
+            mu, logvar = mu.unsqueeze(0), logvar.unsqueeze(0)
         
+        # add p(z) as an expert; regularizes for missing modalities
+        # https://arxiv.org/pdf/1705.10762.pdf
+        # prior_mu, prior_logvar = self.prior((1, mu.size(1), mu.size(2)),
+        #                                      use_cuda=mu.is_cuda)
+        # mu = torch.cat((mu, prior_mu), dim=0)
+        # logvar = torch.cat((logvar, prior_logvar), dim=0)
+        # product of experts to combine gaussians
+        mu, logvar = self.experts(mu, logvar)        
         # reparametrization trick to sample
         z = self.reparametrize(mu, logvar)
         # reconstruct inputs based on that gaussian

### 2017-11-17 mnist/sample.py

@@ -66,6 +66,9 @@ if __name__ == "__main__":
     if not args.condition_on_image and not args.condition_on_text:
         mu = Variable(torch.Tensor([0]))
         std = Variable(torch.Tensor([1]))
+        if args.cuda:
+            mu = mu.cuda()
+            std = std.cuda()
 
     # mode 2: generate conditioned on image
     elif args.condition_on_image and not args.condition_on_text:
@@ -98,7 +101,7 @@ if __name__ == "__main__":
         std = logvar.mul(0.5).exp_()
 
     # sample from uniform gaussian
-    n_latents = vae.encoder.n_latents
+    n_latents = vae.n_latents
     sample = Variable(torch.randn(args.n_samples, n_latents))
     if args.cuda:
         sample = sample.cuda()

### 2017-11-17 mnist/train.py

@@ -64,18 +64,19 @@ def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, tex
     image_BCE, text_BCE = 0, 0
 
     if recon_image is not None and image is not None:
-        image_BCE = lambda_xy * F.binary_cross_entropy(recon_image, image.view(-1, 784), 
-                                                       size_average=False)
+        image_BCE = lambda_xy * F.binary_cross_entropy(recon_image, image.view(-1, 784))
 
     if recon_text is not None and text is not None:
-        text_BCE = lambda_yx * F.nll_loss(recon_text, text, size_average=False)
+        text_BCE = lambda_yx * F.nll_loss(recon_text, text)
 
+    import pdb; pdb.set_trace()
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
     # https://arxiv.org/abs/1312.6114
     # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-    return (image_BCE + text_BCE + KLD / 784) / batch_size
+    KLD /= batch_size * 784  # for each pixel
+    return image_BCE + text_BCE + KLD
 
 
 if __name__ == "__main__":
@@ -137,7 +138,7 @@ if __name__ == "__main__":
             loss_1 = loss_function(mu_1, logvar_1, recon_image=recon_image_1, image=image, 
                                    recon_text=recon_text_1, text=text, lambda_xy=1., lambda_yx=1.)
             loss_2 = loss_function(mu_2, logvar_2, recon_image=recon_image_2, image=image, 
-                                   recon_text=recon_text_2, text=text, lambda_xy=1., lambda_yx=0.5)
+                                   recon_text=recon_text_2, text=text, lambda_xy=1., lambda_yx=0.)
             loss_3 = loss_function(mu_3, logvar_3, recon_image=recon_image_3, image=image, 
                                    recon_text=recon_text_3, text=text, lambda_xy=0., lambda_yx=1.)
             
@@ -178,9 +179,9 @@ if __name__ == "__main__":
             loss_1 = loss_function(mu_1, logvar_1, recon_image=recon_image_1, image=image, 
                                    recon_text=recon_text_1, text=text, lambda_xy=1., lambda_yx=1.)
             loss_2 = loss_function(mu_2, logvar_2, recon_image=recon_image_2, image=image, 
[…]

### 2017-11-17 multimnist/train.py

@@ -73,18 +73,18 @@ def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, tex
     
     if recon_image is not None and image is not None:
         image_BCE = lambda_xy * F.binary_cross_entropy(recon_image.view(-1, 1 * 50 * 50), 
-                                                       image.view(-1, 1 * 50 * 50), size_average=False)
+                                                       image.view(-1, 1 * 50 * 50))
 
     if recon_text is not None and text is not None:
-        text_BCE = lambda_yx * F.nll_loss(recon_text.view(-1, recon_text.size(2)), text.view(-1),
-                                          size_average=False)
+        text_BCE = lambda_yx * F.nll_loss(recon_text.view(-1, recon_text.size(2)), text.view(-1))
 
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
     # https://arxiv.org/abs/1312.6114
     # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-    return (image_BCE + text_BCE + KLD * kl_lambda) / batch_size
+    KLD = KLD / batch_size * kl_lambda
+    return image_BCE + text_BCE + KLD
 
 
 if __name__ == "__main__":

### 2017-11-17 mnist/train.py
KL was too high which was causing blurry samples

@@ -60,7 +60,6 @@ def load_checkpoint(file_path, use_cuda=False):
 
 def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, text=None,
                   lambda_xy=1., lambda_yx=1.):
-    batch_size = mu.size(0)
     image_BCE, text_BCE = 0, 0
 
     if recon_image is not None and image is not None:
@@ -69,13 +68,12 @@ def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, tex
     if recon_text is not None and text is not None:
         text_BCE = lambda_yx * F.nll_loss(recon_text, text)
 
-    import pdb; pdb.set_trace()
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
     # https://arxiv.org/abs/1312.6114
     # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-    KLD /= batch_size * 784  # for each pixel
+    KLD /= args.batch_size * (784 / 3)  # for each pixel
     return image_BCE + text_BCE + KLD
 
 
@@ -138,9 +136,9 @@ if __name__ == "__main__":
             loss_1 = loss_function(mu_1, logvar_1, recon_image=recon_image_1, image=image, 
                                    recon_text=recon_text_1, text=text, lambda_xy=1., lambda_yx=1.)
             loss_2 = loss_function(mu_2, logvar_2, recon_image=recon_image_2, image=image, 
-                                   recon_text=recon_text_2, text=text, lambda_xy=1., lambda_yx=0.)
+                                   recon_text=recon_text_2, text=text, lambda_xy=1., lambda_yx=1.)
             loss_3 = loss_function(mu_3, logvar_3, recon_image=recon_image_3, image=image, 
-                                   recon_text=recon_text_3, text=text, lambda_xy=0., lambda_yx=1.)
+                                   recon_text=recon_text_3, text=text, lambda_xy=1., lambda_yx=1.)
             
             loss = loss_1 + loss_2 + loss_3
             loss.backward()

### 2017-11-17 paper/draft.tex
Rewrite the intro+related works

@@
 For example, one can imagine what a black cat looks like but also prescribe this description to a photo.
-Additionally, available multimodal information is often \textit{sparse}, in the sense there exist many large corpuses of images or text alone but few datasets exist (and are expensive to construct) that contain multiple modalities together.
+Additionally, available multimodal information is often \textit{sparse}.
+There exist many large corpuses of images or text alone but few datasets exist (and are expensive to construct) that contain multiple modalities together.
 Ideally, we would like the generative models we build to extract a joint representation that captures high-level concepts across modalities in a way that is bi-directional and weakly-supervised.
 
-In this work, we propose a multimodal variational autoencoder (MMVAE).
-In our graphical model, we have two modalities $x_{1}$ and $x_{2}$ that are each conditioned on a latent variable $z$, which models the joint distribution, $p(x_{1}, x_{2})$.
-Such a modal is bi-directional since we can reconstruct both modalities from either $x_{1}$ or $x_{2}$.
-We will also show that we can train the MMVAE with sparse $(x_{1}, x_{2})$ pairs, and heavily utilize on a larger set of independent $(x_{1})$ and $(x_{2})$ examples.
-The most significant feature of MMVAEs is that it does not create a separate multimodal encoder $q(z | x_{1}, x_{2})$; instead it shares parameters with the encoders from each modality, $q(z | x_{1})$, $q(z | x_{2})$.
-This has the benefit of producing a simpler model while ensuring that high-dimensional missing modalities do not cause our generated samples to collapse.
+In this work, we propose a multimodal variational autoencoder (MMVAE), an extension of the traditional VAE framework to handle multiple input types.
+The first extension is model the setting where we have both an image, $x$, and a natural text, $y$.
+We then assume a joint generative model of the form $p(x, y, z) = p(z)p(x | z)p(y | z)$ where $p(z)$ is the prior over the latent variable $z$, $p(x | z)$ is our image decoder, and $p(y | z)$ is our text decoder.
+We then further extend the VAE using a novel objective function, which we call the \textit{MMELBO}, for simultaneously training the model from paired data, $D_{p} = \{ (x_{n}, y_{n}) \}$, and unimodal data, $D_{x} = \{ x_{n} \}$, $D_{y} = \{ y_{n} \}$.
+Namely, using MMELBO allows for both evaluating unpaired data at test time and training with sparse paired examples.
+
+MMVAE jointly fits three encoders: $q(z | x, y)$, $q(z | x)$, and $q(z | y)$.
+Doing so, we can project images and natural text into the same latent space, and offer ways to translate between the two (bi-directional).
+For efficiency and simplicity, MMVAEs share parameters between the three encoders.
+The encoders interact via product of experts (POE) \cite{hinton2006training, vedantam2017generative}.
+
+We report experimental results on three datasets.
+The first dataset is MNIST, where labels are stringified to text as the second modality.
+The second dataset is MultiMNIST, where 0 to 4 MNIST digits are placed on a 50x50 canvas.
+Unlike the traditional MultiMNIST, we do not vary the locations of these 0 to 4 digits.
+We again use the stringified label as a the second modality, where the label is the digit class from top-left to bottom-right.
+We briefly discuss introducing randomness into the text by permuting the string.
+The third dataset is the Microsoft COCO Captions dataset \cite{chen2015microsoft}, which contains 330K images and 5 captions per image.
+We show that our model performs competitively on these datasets.
+
+Finally, for each of the three datasets, we investigate how our model performs under weak-supervision by aggressively reducing the number of paired data points we provide to the model.
+We show that MMVAEs do not need a significant number of paired data points to learn a good joint representation across modalities, making this model feasible for real-world tasks i.e.
+caption generation or speech-to-text.
+We also briefly discuss reducing the number of data points of a single modality and its effect of learning.
+
+In summary, the contributions of this paper are as follows: first, we present MMVAEs, introducing the MMELBO objective.
+Second, we show that the product of experts techniques works with two unstructured modalities: images and raw text.
+Lastly, we should that multimodal VAEs are good candidates for weakly supervised learning.
 
@@
 For instance, \citet{ngiam2011multimodal} trained deep autoencoders on audio and video input and found that the bimodal representations were better than any single equivalent.
-More recently, VAEs \cite{kingma2013auto, kingma2014semi} have been used to train models in high-dimensional multimodal settings.
-For example, conditional VAEs \cite{sohn2015learning} maximize a conditional log-likleihood by variational methods where often one modality is conditioned on another i.e.
-handwriting digits and labels, faces and attributes, or captioning.
-Notably, conditional VAEs (CVAE) are not bi-directional.
+More recently, VAEs \cite{kingma2013auto, kingma2014semi} have been used to train conditional generative models of the form $p(y | x)$ in high-dimensional bimodal settings.
+Here, $y$ is often a class label, a sentence, or another image depending on the task.
+For example, conditional VAEs \cite{sohn2015learning}, and conditional multimodal autoencoders (CMMA) \cite{pandey2017variational} maximize a conditional log-likelihood where often one modality is conditioned on another i.e.
+handwriting digits and labels, faces and attributes, etc.
+Notably, these CVAEs and CMMAs are not bi-directional.
+We are more interested in learning a shared latent space of $p(x, y)$ where we can condition interchangeably.
 
-\citet{pandey2017variational} proposed the conditional multimodal autoencoder (CMMA), which also maximizes a conditional log-likelihood.
-CMMA explicitly connects the conditional variable the latent variable, where CVAE does not.
-Moreover, it also constrains the latent representation to be close to the joint representation across modalities.
-However, CMMA is still not bi-directional.
+Recently, there have been several works on such joint models, many of which share commonalities to our proposed model.
+For example, the BiVCCA employs a similar VAE model with the objective shown below.
+It includes two inference networks, one for each modality.
+We note that MMVAE uses a third multimodal objective; we show in our experiments that doing so is important in learning a useful latent space.
+The BiVCCA loss encourages proper reconstruction of both $x$ and $y$ given only one; $\mu$ acts as a weight to prioritize different modalities.
 
-Finally, \citet{suzuki2016joint} proposed a joint multimodal VAE (JMVAE).
-This is similar to our work in that it is bi-directional, handles missing modalities, and models the joint distribution.
-Our proposed model (MMVAE) and JMVAE share the same graphical notation, but differ in our approach.
-Crucially, JMVAE trains separate encoders for each modality and an additional one for the multimodal case.
-It then includes a KL divergence term \textit{per modality} in the loss.
-In constrast, MMVAE does not have an additional encoder for the multimodal case and does not have additional KL terms.
-It does so by sharing encoder parameters to learn the guide distributions: $q_{\phi_{x_{1}}}(z|x_{1})$, $q_{\phi_{x_{2}}}(z|x_{2})$, $q_{\phi}(z|x_{1}, x_{2})$
+\begin{equation}
+    J(x, y, \theta, \phi) = \mu (E_{q_{\phi_{x}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{x}}(z|x)), p_{\theta}(z)) + (1 - \mu) (E_{q_{\phi_{y}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{y}}(z|y)), p_{\theta}(z))
+\end{equation}
+
+\citet{suzuki2016joint} then build onto the BiVCCA objective with their JMVAE model, that adds an additional multimodal encoder $q(z|x,y)$ and two additional KL divergence terms to bring $q(z|x,y)$ close to $q(z|x)$ and $q(z|y)$.
+More explicitly, their objective is the following:
+
+\begin{equation}
+    J(x, y, \theta, \phi) = \textup{elbo}_{1, \lambda, 1}(x, y, \theta, \phi) - \alpha[\textup{KL}(q_{\phi}(z|x,y), q_{\phi_{y}}(z|y)) + \textup{KL}(q_{\phi}(z|x,y), q_{\phi_{x}}(z|x))]
+\end{equation}
+
+As \citet{vedantam2017generative} note, forcing $q_{\phi(z|y)}$ to be close to $q_{\phi(z|x, y)}$ is somewhat unintuitive but if you rewrite $q_{\phi(z|x, y)}$ as $q_{\phi}^{\textup{avg}}(z|y)$, they show that it has the benefit of $q_{\phi(z|y)}$ covering the embeddings of all $x$ associated with $y$.
+\citet{vedantam2017generative} go on to introduce their own multimodal VAE objective, the \textit{TELBO} to achieve a similar effect with a simpler design.
+The TELBO is defined as
+
+\begin{equation}
+    J(x, y, \theta_{x}, \theta_{y}, \phi, \phi_{x}, \phi_{y}) = \textup{elbo}_{1, \lambda, 1}(x, y, \theta_{x}, \theta_{y}, \phi) + \textup{elbo}_{1, 1}(x, \theta_{x}, \phi_{x}) + \textup{elbo}_{\gamma, 1}(y, \theta_{y}, \phi_{y})
+\end{equation}
+
+where $\lambda, \gamma$ are hyperparameters of the ELBO.
+See model section for notation.
+The authors report training this model in two stages: first, using only aligned data, train on first ELBO term alone.
+Then, freezing the decoders $p_{\theta_{x}}(x|z)$ and $p_{\theta_{y}}(y|z)$, train on the last two ELBO terms above.
+The TELBO is very similar to our proposed objective (MMELBO).
+We denote two main differences.
+First, in TELBO, when the input is a single modality, the objective decomposes to the traditional ELBO, where it penalizes differences between the reconstruction and original of that modality only.
+In MMELBO, all three ELBO terms include reconstruction costs for all modalities.
+For example, given an unpaired image $x$, MELBO encourages its latent representation to decode to both a sensible image and text.
+We later show that this change is important in produce sharper samples.
+Second, training MMELBO is done concurrently in a single stage.
+We show paired and unpaired examples together and share encoder and decoder parameters across the three ELBO terms.
 

### 2017-11-17 paper/compile.sh

@@ -1,0 +1,7 @@
+#!/bin/bash
+
+bibtex draft
+bibtex draft
+pdflatex draft
+bibtex draft
+pdflatex draft

### 2017-11-17 paper/draft.tex
Finished rewriting everything -- onto the experiments part now

@@
 \usepackage{microtype}
-\usepackage{nips_2017}
+\usepackage[final]{nips_2017}
 \usepackage{graphicx}
@@
 \begin{abstract}
-TODO
+Last thing TODO.
 \end{abstract}
@@
 We then assume a joint generative model of the form $p(x, y, z) = p(z)p(x | z)p(y | z)$ where $p(z)$ is the prior over the latent variable $z$, $p(x | z)$ is our image decoder, and $p(y | z)$ is our text decoder.
-We then further extend the VAE using a novel objective function, which we call the \textit{MMELBO}, for simultaneously training the model from paired data, $D_{p} = \{ (x_{n}, y_{n}) \}$, and unimodal data, $D_{x} = \{ x_{n} \}$, $D_{y} = \{ y_{n} \}$.
-Namely, using MMELBO allows for both evaluating unpaired data at test time and training with sparse paired examples.
+We then further extend the VAE using a novel objective function, which we call the multimodal ELBO (\textit{MELBO}), for simultaneously training the model from paired data, $D_{p} = \{ (x_{n}, y_{n}) \}$, and unimodal data, $D_{x} = \{ x_{n} \}$, $D_{y} = \{ y_{n} \}$.
+Namely, using MELBO allows for both evaluating unpaired data at test time and training with sparse paired examples.
 
@@
 
-In summary, the contributions of this paper are as follows: first, we present MMVAEs, introducing the MMELBO objective.
+In summary, the contributions of this paper are as follows: first, we present MMVAEs, introducing the MELBO objective.
 Second, we show that the product of experts techniques works with two unstructured modalities: images and raw text.
@@
 Then, freezing the decoders $p_{\theta_{x}}(x|z)$ and $p_{\theta_{y}}(y|z)$, train on the last two ELBO terms above.
-The TELBO is very similar to our proposed objective (MMELBO).
-We denote two main differences.
+The TELBO is very similar to our proposed objective (MELBO).
+We denote three main differences.
 First, in TELBO, when the input is a single modality, the objective decomposes to the traditional ELBO, where it penalizes differences between the reconstruction and original of that modality only.
-In MMELBO, all three ELBO terms include reconstruction costs for all modalities.
+In MELBO, all three ELBO terms include reconstruction costs for all modalities.
 For example, given an unpaired image $x$, MELBO encourages its latent representation to decode to both a sensible image and text.
 We later show that this change is important in produce sharper samples.
-Second, training MMELBO is done concurrently in a single stage.
+Second, training MELBO is done concurrently in a single stage.
 We show paired and unpaired examples together and share encoder and decoder parameters across the three ELBO terms.
+Thirdly, \citet{vedantam2017generative} only test on images and one-hotted attributes in their experiments.
+In this work, we will explore a more difficult second modality, unstructured text.
 
 \section{Methods}
-This section first briefly goes over VAEs, and product of Gaussians (PoG), then details a new multimodal VAE (MMVAE).
+This section first briefly goes over standard VAEs, and product of Gaussians (PoG), then details a new multimodal VAE (MMVAE).
+See supplement for details of the MLP, CNN, and RNN architectures used as encoders and decoders.
 
 \subsection{Variational Autoencoders}
-Variational autoencoders (VAE) model a generative process as $z \sim \mathcal{N}(0, I)$ and $x \sim p_{\theta}(x | z)$ where $x$ is an observed variable, $z$ is a latent variable, and $\theta$ are the model parameters.
-VAE wish to maximize the marginal distribution $p(x) = \int p_{\theta}(x | z)p(z)dx$, which is intractable.
-Instead, we maximize the lower bound of the marginal distribution:
+A variational autoencoder (VAE) \cite{kingma2013auto} is a latent generative model of the form $p_{\theta}(x, z) = p_{\theta}(z)p_{\theta}(x|z)$ where $p_{\theta}(z)$ is usually a Gaussian prior, and $p_{\theta}(x|z)$ is the decoder, representing the likelihood.
+Training takes the form of fitting an encoder network, $q_{\phi}(z|x)$ to maximize the evidence lower bound (ELBO), defined below:
 
-\begin{equation}
-    \textup{log } p(x) \geq -D_{KL}(q_{\phi}(z | x) || p(z)) + E_{q_{\phi}}[\textup{log } p_{\theta}(x | z)]
+\begin{equation}\
+    \textup{elbo}_{\lambda, \beta}(x, \theta, \phi) = E_{q_{\phi}(z|x, \phi)}[\lambda \textup{ log}(p_{\theta}(x|z))] - \beta \textup{ KL}(q_{\phi}(z|x), p_{\theta}(z))
 \end{equation}
 
-where $q_{\phi}(z | x)$ is a \textit{guide distribution} that approximates the posterior $p(z | x)$; $\phi$ are the parameters of $q$.
-In the text, we call $p_{\theta}(x | z)$ a decoder and $q_{\theta}(z | x)$ an encoder.
-In practice, $q$ is set to the familiy of Gaussian distributions, where optimization amounts to learning $\phi=\{\mu, \sigma\}$.
-The standard reparametrization trick is employed to sample $z$.
+$\textup{KL}(p, q)$ is the Kullback Leibler divergence between distributions $p$ and $q$; $\beta$ and $\lambda$ are weights for the different terms in the ELBO.
+In practice, increasing $\beta$ encourages a better latent embedding but often leads to blurry samples.
 
@@
 
-We wish to model the interaction between our modalities as a product of experts.
-Fortunately, because our guide distributions are Gaussian, MMVAEs take advantage of the fact the product of two Gaussian distributions is itself a Gaussian distribution \cite{cao2014generalized}.
-The mean and covariance of the resulting Gaussian distribution is:
+We wish to model the interaction between our modalities as a product of experts \cite{hinton2006training}.
+Fortunately, because our guide distributions are Gaussian, the product of two Gaussian distributions is itself Gaussian \cite{cao2014generalized}.
+The resulting Gaussian distribution has mean $mu = (\sum_{i} \mu_{i}\Sigma^{-1}_{i})(\sum_{i}\Sigma^{-1}_{i})^{-1}$ and covariance $\Sigma = (\sum_{i} \Sigma^{-1}_{i})^{-1}$, where $\mu_{i}$, $\Sigma_{i}$ is the mean and covariance of the $i$-th Gaussian expert.
+In \cite{hinton2006training}, the prior $p_{\theta}(z)$ is included as an expert for regularization; we omit this term, having found it unstable in practice.
+
+\subsection{Multimodal Variational Autoencoders}
+
+A multimodal variational autoencoder is a model of the form $p_{\theta}(x, y, z) = p_{\theta}(z)p_{\theta}(x|z)p_{\theta}(y|z)$ where $x$ and $y$ are two modalities, i.e.
+images and text, $z$ is the latent variable, $p_{\theta}(x|z)$ is a image decoder, and $p_{\theta}(y|z)$ is a text decoder.
+The training objective is as in standard VAE, with the exception that we expand the evidence lower bound definition to include dual modalities.
 
 \begin{equation}
-    \mu = (\sum_{i} \mu_{i}\Sigma^{-1}_{i})(\sum_{i}\Sigma^{-1}_{i})^{-1}
-\end{equation}
-\begin{equation}
-    \Sigma = (\sum_{i} \Sigma^{-1}_{i})^{-1}
+    \textup{elbo}_{\lambda_{x}, \lambda_{y}, \beta}(x, \theta_{x}, \theta_{y}, \phi) = E_{q_{\phi}(z|x, \phi)}[\lambda_{x} \textup{ log}(p_{\theta_{x}}(x|z)) + \lambda_{y} \textup{ log}(p_{\theta_{y}}(y|z))] - \beta \textup{ KL}(q_{\phi}(z|x, y), p_{\theta}(z))
 \end{equation}
 
-where $\mu_{i}$, $\Sigma_{i}$ is the mean and covariance of the $i$-th Gaussian expert.
+Note that we have additional hyperparameters, $\lambda_{x}$ and $\lambda_{y}$ that allow us to prioritize one modality over the other.
+This expanded ELBO is used for paired and unpaired examples.
+Combining these three terms gives us the Multimodal ELBO (MELBO) objective:
 
-\subsection{Multimodal Variational Autoencoders}
-Given a dataset $(X_{1}, X_{2}) = \{ (x_{1,1}, x_{2,1}), ..., (x_{1,N}, x_{2,N}) \}$, where $x_{1}$ and $x_{2}$ are two modalities of possibly different dimensions.
-We assume that $x_{1}$ and $x_{2}$ are conditionally independent on the latent state $z$.
-Figure \ref{fig:diagram:graph} shows the graphical model of the generative process: $z \sim p(z)$ and $x_{1}, x_{2} \sim p(x_{1}, x_{2} | z)$.
+\begin{equation}
+    J(x, y, \theta_{x}, \theta_{y}, \phi) = \textup{elbo}_{\lambda_{x}^{1}, \lambda_{y}^{1}, \beta^{1}}(x, y, \theta_{x}, \theta_{y}, \phi) + \textup{elbo}_{\lambda_{x}^{2}, \lambda_{y}^{2}, \beta^{2}}(x, \theta_{x}, \phi) + \textup{elbo}_{\lambda_{x}^{3}, \lambda_{y}^{3}, \beta^{3}}(y, \theta_{y}, \phi)
+\end{equation}
 
-We can estimate the lower bound of the log-likelihood $\textup{log }p(x_{1}, x_{2})$ as follows:
-
-\begin{multline}
-\textup{log }p(x_{1}, x_{2}) \geq -D_{KL}(q_{\phi}(z | x_{1}, x_{2}) || p(z)) \\ + E_{q_{\phi}(z | x_{1}, x_{2})}[\textup{log } p_{\theta_{x_{1}}}(x_{1} | z)] + E_{q_{\phi}(z | x_{1}, x_{2})}[\textup{log } p_{\theta_{x_{2}}}(x_{2} | z)]
-\end{multline}
-
-The equation above has two reconstruction terms, one for each modality.
-As with VAEs, we can parametrize the encoder and decoder as deep neural networks and use stochastic gradient descent to optimize the parameters, $\theta_{x_{1}}$, $\theta_{x_{2}}$, $\phi$.
-Because each modality learns a different feature representation, each modality has its own encoder and decoder.
-
-In MMVAE, it is clear that we can compute the marginal and conditional distribution in both directions: given either $x_{1}$ or $x_{2}$ we can reconstruct both modalities.
-Additionally, we can easily extend MMVAE to handle more than two modalities, $p(x_{1}, x_{2}, ..., x_{n})$ using the same framework.
+The $\lambda$s scale the individual log likelihood terms and the $\beta$s scale the KL divergence terms.
+We set these parameters using a validation set on a per-dataset level.
+During training, we shown both paired and unpaired data; and we train the three encoders and decoders together with shared weights.
+Regardless of paired or unpaired examples, our objective requires both modalities to be reconstructed, sharing parameters for all decoders.
+We found that this is crucial to learning a good latent space between images and text.
+This is, to the best of our knowledge, a novel technique.
+Note that if $\lambda_{y}^{2}$ and $\lambda_{x}^{3}$ are set 0, the MELBO is equivalient to the TELBO \cite{hinton2006training}.
 
@@
 
+\subsection{Missing Modalities} At training and test time, we can handle missing attributes by using a product of Gaussian experts, as in \cite{hinton2006training}.
+The inference network takes the form: $q(z|y_{1:k})$.
+Missing modalities then translates to fewer experts; training proceeds untouched.
+
 \subsection{Weakly-Supervised Training}
-In most circumstances, paired modal data is both expensive and scarce while unimodal data, i.e.
-images or text is more easily available.
+In most circumstances, paired data is both expensive and scarce while unimodal data, i.e.
+images or text is more easily available in the natural world.
 The design of MMVAE allows for bootstrapping under sparse labels.
-During training, we consider input pairs of 3 types: the 1st modality alone,  the 2nd modality alone, or two separate modalities.
+During training, we consider input pairs of three types: the first modality alone, the second modality alone, or two paired modalities.
 Figure \ref{fig:diagram:modelv2}, \ref{fig:diagram:modelv1}, \ref{fig:diagram:modelv3} enumerate these three cases.
-In the case of a single modality, training is identical to VAE; the encoder $E_{i}(x_{i})$ returns $\mu_{i}$, $\sigma_{i}$, which is used to sample $z$ and decode.
-In the case of two modalities, each encoder returns $\mu_{i}$, $\sigma_{i}$, which we can combine via PoG to return the parameters of the joint Gaussian distribution.
-This is then used to sample $z$ and decode as in the traditional VAE.
-Crucially, we do not need additional loss terms or separate networks to handle unimodal and multimodal cases; the only change required is an additional PoG operation.
-In practice, we find that we only need a limited amount of paired multimodal examples in order to learn the cross-modal relations.
+If given a paired example, use the full MELBO loss.
+If given only the first modality, decompose the objective to $J(x, \theta_{x}, \phi) = \textup{elbo}_{\lambda_{x}, 0, \beta}(x, \theta_{x}, \phi)$.
+If given only the second modality, decompose the objective to $J(x, \theta_{y}, \phi) = \textup{elbo}_{0, \lambda_{y}, \beta}(x, \theta_{y}, \phi)$.
+In other words, when only a single modality is present, use the standard VAE objective.
+In practice, we find that we only need a limited amount of paired multimodal examples in order to learn the latent space for our datasets.
+
+\section{Experimental Results}
+
+In this section, we fit the MMVAE to three datasets (MNIST, MultiMNIST, and COCO-Captions) using the MELBO objective.
+We measure the quality by approximating the marginal log likelihood with a lower bound, and through qualitative analysis of generated samples.
+We also perform a series of weak-supervision experiments on each dataset, progressively lowering the fraction of paired examples, and unpaired examples and observing performance.
+
+\subsection{MNIST}
+\subsection{MultiMNIST}
+\subsection{COCO-Captions}
 

### 2017-11-17 paper/draft.tex
Added descriptions about datasets

@@
 
-We report experimental results on three datasets.
+We report experimental results on four datasets.
 The first dataset is MNIST, where labels are stringified to text as the second modality.
@@
 We again use the stringified label as a the second modality, where the label is the digit class from top-left to bottom-right.
-We briefly discuss introducing randomness into the text by permuting the string.
 The third dataset is the Microsoft COCO Captions dataset \cite{chen2015microsoft}, which contains 330K images and 5 captions per image.
+The fourth dataset is the LibriSpeech dataset, which contains X waveforms and their text transcriptions.
 We show that our model performs competitively on these datasets.
 
-Finally, for each of the three datasets, we investigate how our model performs under weak-supervision by aggressively reducing the number of paired data points we provide to the model.
+Finally, for each of the four datasets, we investigate how our model performs under weak-supervision by aggressively reducing the number of paired data points we provide to the model.
 We show that MMVAEs do not need a significant number of paired data points to learn a good joint representation across modalities, making this model feasible for real-world tasks i.e.
@@
 Fortunately, because our guide distributions are Gaussian, the product of two Gaussian distributions is itself Gaussian \cite{cao2014generalized}.
-The resulting Gaussian distribution has mean $mu = (\sum_{i} \mu_{i}\Sigma^{-1}_{i})(\sum_{i}\Sigma^{-1}_{i})^{-1}$ and covariance $\Sigma = (\sum_{i} \Sigma^{-1}_{i})^{-1}$, where $\mu_{i}$, $\Sigma_{i}$ is the mean and covariance of the $i$-th Gaussian expert.
+The resulting Gaussian distribution has mean $\mu = (\sum_{i} \mu_{i}\Sigma^{-1}_{i})(\sum_{i}\Sigma^{-1}_{i})^{-1}$ and covariance $\Sigma = (\sum_{i} \Sigma^{-1}_{i})^{-1}$, where $\mu_{i}$, $\Sigma_{i}$ is the mean and covariance of the $i$-th Gaussian expert.
 In \cite{hinton2006training}, the prior $p_{\theta}(z)$ is included as an expert for regularization; we omit this term, having found it unstable in practice.
@@
 
-In this section, we fit the MMVAE to three datasets (MNIST, MultiMNIST, and COCO-Captions) using the MELBO objective.
+In this section, we fit the MMVAE to four datasets (MNIST, MultiMNIST, COCO-Captions, LibriSpeech) using the MELBO objective.
 We measure the quality by approximating the marginal log likelihood with a lower bound, and through qualitative analysis of generated samples.
@@
 \subsection{MNIST}
+
+This is the original MNIST dataset except the labels are casted as strings and treated as a separate modality. We use 50k for training, 10k validation, 10k testing.
+
+\paragraph{Models} We train the MMVAE model using the MELBO objective.
+We use Adam for optimization with default learning rate of $1e-3$, a minibatch size of 128, 20 latent dimensions, and train for 20 epochs.
+
+For the image encoder, $p(x|z)$ and decoder, $q(z|x)$ we use a three layer MLP with batch normalization and ReLU nonlinearity.
+For the text encoder $p(y|z)$ and decoder $q(z|y)$, we use a continuous embedding of size 10, and a two layer MLP, again with batch normalization and ReLU.
+$q(z|y)$ additionally has a softmax to compute probabilities of each word in our vocabulary.
+The joint decoder, $q(z|x,y)$ is the same as $q(z|y)$, $q(x|y)$ since training shares parameters.
+
+For MELBO parameters, we set $\beta^{i} = 0.004$, and $\lambda_{x}^{i} = \lambda_{y}^{i} = 1, i \in \{1,2,3\}$.
+
+\paragraph{Evaluation} We first measure the marginal log likelihoods, $\textup{log}(p_{\textup{test}}(x))$, $\textup{log}(p_{\textup{test}}(y))$ by computing the ELBO with 1000 particles for importance sampling on 10k test set.
+For each test example, we can compute the ELBO using both modalities, just images, or just text.
+This results in 6 marginal probabilities, shown in table.
+Since MNIST is a prediction task, we can measure the predictive accuracy of our model by generating a text sample conditioned on a image as the label for that image.
+In table, we measure these nine metrics for a variety of MELBO variations.
+
+\paragraph{Weak Supervision} We measure the performance of our model on predicting the correct MNIST label as the number of paired examples we show to the model approach 0.
+We also measure the same performance as the number of unpaired images/text we show approach 0.
+
 \subsection{MultiMNIST}
+
+This is a variant of the original MNIST dataset where between 0 and 4 MNIST digits are composed together on a 50x50 canvas.
+Given a MultiMNIST image, we generate the paired text example by concatenating the digit classes from top-left to bottom-right.
+Notably, this is a much more difficult text task than our MNIST dataset due to variable text length and a much larger space of acceptable words.
+We again use 50k for training, 10k for validation, 10k for testing.
+
+\paragraph{Models} Similar training parameters as in MNIST except we use a minibatch size of 64, 100 latent dimensions, and train for 40 epochs due to larger encoder/decoder models.
+
+For the image encoder and decoder, we use the DCGAN architecture from.
+For the text encoder, we use a continuous embedding layer of size 10, followed by a bidirectional GRU with $50$ hidden units, and a single fully connected layer to project into our latent dimension size.
+For the text decoder, we generate a single character at a time.
+We include a start token and a stop token in our vocabulary.
+Generating text is as follows: provided a start token, get its continuous embedding, feed it through a fully connected layer, then a unidirectional GRU, followed by another fully connected layer and a softmax over the 10 character space.
+We sample a new character and repeat the process until generating a stop token.
+See supplement for more details.
+
+For MELBO parameters, we anneal $\beta^{i}$ every 5 epochs with the following schedule: 1e-5, 1e-4, 1e-3, 1e-2, 1e-1.
+We set $\lambda_{y}^{2} = 0.5$ and $\lambda_{x}^{1} = 0.1$.
+All other $\lambda$s are set to 1.
+
+\paragraph{Evaluation and Weak Supervision} Setup is identical to the MNIST setting except we are measuring (1) that the length of the generated text matches the ground truth, (2) accuracy of matching characters i.e.
+2845 and 3845 would be 75\% accuracy, and (3) accuracy of matching sequences i.e.
+2845 and 3845 would be incorrect.
+
 \subsection{COCO-Captions}
+
+\subsection{LibriSpeech}
+TODO.
 

### 2017-11-17 paper/draft.tex
Done with words -- need figures and tables

@@
 
+The Microsoft COCO-Captions dataset contains 330k natural images, each tagged with 5 captions.
+Images are all rescaled to 64x64 and with a 64x64 center crop.
+When training, every epoch, we randomly sample a caption from the 5 available captions per image.
+At test time, we use the first caption.
+We use X for training, X for validation, and X for testing.
+Both modalities are more difficult than the MultiMNIST setting.
+
+\paragraph{Models} Identical training parameters to MultiMNIST.
+VAEs suffer from blurry samples when training on natural images.
+To address this, we replace the DCGAN decoder with a PixelVAE decoder.
+For the text encoder and decoders, we increase the hidden units to 500.
+Instead of learning our own continuous embedding, we use pre-trained GloVe embeddings for each word in our caption to support the large vocabulary size.
+Generation remains the same with a start and stop token.
+Identical MELBO parameters to MultiMNIST.
+
+\paragraph{Evaluation and Weak Supervision} There is no predictive task for COCO-Captions. Instead we measure the marginal loglikelihood for weak supervision tests.
+
 \subsection{LibriSpeech}
+TODO.
+
+\section{Discussion}
+TODO.
+
+\section{Conclusions and Future Work}
 TODO.

### 2017-11-18 paper/draft.tex
Added images + placeholder tables for MNIST... need to do for MultiMNIST and then call it a day

@@
 \usepackage{natbib}
+\usepackage{breqn}
 
@@
 
-\begin{equation}
-    J(x, y, \theta, \phi) = \mu (E_{q_{\phi_{x}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{x}}(z|x)), p_{\theta}(z)) + (1 - \mu) (E_{q_{\phi_{y}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{y}}(z|y)), p_{\theta}(z))
-\end{equation}
+\begin{multline}
+    J(x, y, \theta, \phi) = \mu (E_{q_{\phi_{x}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{x}}(z|x)), p_{\theta}(z)) \\ + (1 - \mu) (E_{q_{\phi_{y}}(z|x)}[\textup{log}(p_{\theta_{x}}(x|z)) + \lambda \textup{ log}(p_{\theta_{y}}(y|z))] - \textup{KL}(q_{\theta_{y}}(z|y)), p_{\theta}(z))
+\end{multline}
 
@@
 
-\begin{equation}
-    \textup{elbo}_{\lambda_{x}, \lambda_{y}, \beta}(x, \theta_{x}, \theta_{y}, \phi) = E_{q_{\phi}(z|x, \phi)}[\lambda_{x} \textup{ log}(p_{\theta_{x}}(x|z)) + \lambda_{y} \textup{ log}(p_{\theta_{y}}(y|z))] - \beta \textup{ KL}(q_{\phi}(z|x, y), p_{\theta}(z))
-\end{equation}
+\begin{multline}
+    \textup{elbo}_{\lambda_{x}, \lambda_{y}, \beta}(x, \theta_{x}, \theta_{y}, \phi) = E_{q_{\phi}(z|x, \phi)}[\lambda_{x} \textup{ log}(p_{\theta_{x}}(x|z)) + \lambda_{y} \textup{ log}(p_{\theta_{y}}(y|z))] \\ - \beta \textup{ KL}(q_{\phi}(z|x, y), p_{\theta}(z))
+\end{multline}
 
@@
 
-\begin{equation}
-    J(x, y, \theta_{x}, \theta_{y}, \phi) = \textup{elbo}_{\lambda_{x}^{1}, \lambda_{y}^{1}, \beta^{1}}(x, y, \theta_{x}, \theta_{y}, \phi) + \textup{elbo}_{\lambda_{x}^{2}, \lambda_{y}^{2}, \beta^{2}}(x, \theta_{x}, \phi) + \textup{elbo}_{\lambda_{x}^{3}, \lambda_{y}^{3}, \beta^{3}}(y, \theta_{y}, \phi)
-\end{equation}
+\begin{multline}
+    J(x, y, \theta_{x}, \theta_{y}, \phi) = \textup{elbo}_{\lambda_{x}^{1}, \lambda_{y}^{1}, \beta^{1}}(x, y, \theta_{x}, \theta_{y}, \phi) \\ + \textup{elbo}_{\lambda_{x}^{2}, \lambda_{y}^{2}, \beta^{2}}(x, \theta_{x}, \phi) + \textup{elbo}_{\lambda_{x}^{3}, \lambda_{y}^{3}, \beta^{3}}(y, \theta_{y}, \phi)
+\end{multline}
 
@@
 
+\begin{table}[!h]
+    \centering
+    \begin{tabular}{ l | c | c }
+        Model (MNIST) & $\leq \textup{log }p(x)$ & $\leq \textup{log }p(y)$ \\
+        \hline
+        VAE & & - \\
+        BiVCCA & &  \\
+        JMVAE & & \\
+        MMVAE (Joint) & & \\
+        MMVAE (Image-only) & & \\
+        MMVAE (Text-only) & & \\
+        MMVAE (Anneal KL) & & \\
+        MMVAE (With Universal Expert) & & \\
+        MMVAE ($\lambda_{y}^{1} = \lambda_{x}^{1} = 0$) & & \\
+        MMVAE ($\lambda_{y}^{2} = \lambda_{x}^{3} = 0$) & & \\
+        MMVAE ($\lambda_{x}^{2} = \lambda_{y}^{3} = 0$) & & \\
+        MMVAE ($\lambda_{x}^{2} = \lambda_{y}^{2} = \lambda_{x}^{3} = \lambda_{y}^{3} = 0$) & & \\
+        \newline
+    \end{tabular}
+    \caption{TODO}
+    \label{table:mnist:marginal}
+\end{table}
+
+\begin{figure}[!h]
+\centering
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{mnist_mmvae_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{mnist_mmvae_condition_on_image_text_9_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{mnist_mmvae_condition_on_image_9_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{mnist_mmvae_condition_on_text_9_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \caption{\emph{MNIST Dataset}: MMVAE is bidirectional, allowing us to try samples unconditionally, or conditioned on modalit(ies) taking a particular value.
+We note that samples conditioned on both image and text are the most sharp and have the lowest variance.
+As expected, reconstructed images from text alone are the blurriest.
+(a) Image samples from $z \sim N(0, 1)$.
+(b) Image samples from $p(z|x=9,y=9)$.
+(c) Image samples from $p(z|x=9)$.
+(d) Image samples from $p(z|y=9)$.}
+    \label{fig:mnist:samples}
+\end{figure}
+
+\begin{table}[!h]
+    \centering
+    \begin{tabular}{ l | c | c | c | c | c | c | c | c | c | c | c}
+        (a) & 5 & 7 & 1 & 3 & 1 & 6 & 3 & 7 & 5 & 6 & 0 \\
+        \hline
+        (b) & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 \\
+        \hline
+        (c) & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 \\
+        \hline
+        (d) & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 & 9 \\
+        \newline
+    \end{tabular}
+    \caption{\emph{MNIST Dataset}: Generated text samples from MMVAE trained on MNIST.
+(a) Text samples from $z \sim N(0, 1)$.
+(b) Text samples from $p(z|x=9,y=9)$.
+(c) Text samples from $p(z|x=9)$.
+(d) Text samples from $p(z|y=9)$.
+From (b)-(d), we note that conditioning on an image produces the expected text samples.}
+    \label{table:mnist:samples}
+\end{table}
+
+\begin{figure}[!h]
+\centering
+    \begin{subfigure}[b]{.32\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{mnist_mmvae_pca_manifold.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.32\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{mnist_mmvae_joint_weak_supervision.png}
+        \caption{}
+    \end{subfigure}
+    \caption{\emph{MNIST Dataset}: (a) Manifold split by MNIST class; this was generated using the top two principal components from 10k latent samples of dimension 20.
+We can see that each of the classes form a distinct cluster, suggesting that the latent space is well-defined.
+(b) The effects of weak supervision on the MNIST prediction task where paired data is only shown a fraction of the time; with 25\% supervision, the model reaches 80\% accuracy.}
+    \label{fig:mnist:weaksupervision}
+\end{figure}
+
 \subsection{MultiMNIST}
@@
 
+\begin{table}[!h]
+    \centering
+    \begin{tabular}{ l | c | c }
+        Model (MultiMNIST) & $\leq \textup{log }p(x)$ & $\leq \textup{log }p(y)$ \\
+        \hline
+        VAE & & - \\
+        BiVCCA & &  \\
+        JMVAE & & \\
+        MMVAE (Joint) & & \\
+        MMVAE (Image-only) & & \\
+        MMVAE (Text-only) & & \\
+        MMVAE (Anneal KL) & & \\
+        MMVAE (With Universal Expert) & & \\
+        MMVAE ($\lambda_{y}^{1} = \lambda_{x}^{1} = 0$) & & \\
+        MMVAE ($\lambda_{y}^{2} = \lambda_{x}^{3} = 0$) & & \\
+        MMVAE ($\lambda_{x}^{2} = \lambda_{y}^{3} = 0$) & & \\
+        MMVAE ($\lambda_{x}^{2} = \lambda_{y}^{2} = \lambda_{x}^{3} = \lambda_{y}^{3} = 0$) & & \\
+        \newline
+    \end{tabular}
+    \caption{TODO}
+    \label{table:multimnist:marginal}
+\end{table}
+
 \subsection{COCO-Captions}
@@
 
-\paragraph{Evaluation and Weak Supervision} There is no predictive task for COCO-Captions. Instead we measure the marginal loglikelihood for weak supervision tests.
+\paragraph{Evaluation and Weak Supervision} There is no predictive task for COCO-Captions. Instead we measure the marginal log likelihood for weak supervision tests.
+
+\begin{table}[!h]
+    \centering
+    \begin{tabular}{ l | c | c }
+        Model (COCO-Caption) & $\leq \textup{log }p(x)$ & $\leq \textup{log }p(y)$ \\
+        \hline
+        VAE & & - \\
+        BiVCCA & &  \\
+        JMVAE & & \\
+        MMVAE (Joint) & & \\
+        MMVAE (Image-only) & & \\
+        MMVAE (Text-only) & & \\
+        \newline
+    \end{tabular}
+    \caption{TODO}
+    \label{table:mnist:marginal}
+\end{table}
+
+\begin{figure}[!h]
+    \centering
+    \includegraphics[width=.5\linewidth]{coco_mmvae_image_sample.png}
+    \caption{This is a work in progress.
+These samples were generated with a DCGAN encoder and decoder, hence they are very blurry.
+However, we can make out coherent structure if we squint.
+I'm hoping that with a PixelCNN decoder (PixelVAE), we will get sharper samples.}
+    \label{fig:WIP:coco:samples}
+\end{figure}
 

### 2017-11-18 paper/draft.tex
Updates to draft.tex

@@
 We can see that each of the classes form a distinct cluster, suggesting that the latent space is well-defined.
-(b) The effects of weak supervision on the MNIST prediction task where paired data is only shown a fraction of the time; with 25\% supervision, the model reaches 80\% accuracy.}
+(b) The effects of weak supervision on the MNIST prediction task where paired data is only shown a fraction of the time; with 25\% supervision, the model reaches 80\% accuracy.
+TODO: add weak supervision tests for individual modalities.}
     \label{fig:mnist:weaksupervision}
@@
 
-For the image encoder and decoder, we use the DCGAN architecture from.
+For the image encoder and decoder, we use the DCGAN architecture from \cite{radford2015unsupervised}.
 For the text encoder, we use a continuous embedding layer of size 10, followed by a bidirectional GRU with $50$ hidden units, and a single fully connected layer to project into our latent dimension size.
@@
 
+\begin{figure}[!h]
+\centering
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_image_text_1773_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_text_1773_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \caption{\emph{MultiMNIST Dataset}: (a) Image samples from $z \sim N(0, 1)$.
+(b) Image samples from $p(z|x=1773,y=1773)$.
+(c) Image samples from $p(z|y=1773)$.
+The quality of the samples from (b) and (c) are almost identical.}
+    \label{fig:multimnist:samples}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+    \begin{subfigure}[b]{.32\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_joint_weak_supervision.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.32\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_modal_char_weak_supervision.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.32\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_modal_length_weak_supervision.png}
+        \caption{}
+    \end{subfigure}
+    \caption{\emph{MultiMNIST Dataset}: a series of weak supervision experiments.
+(a) Effect of reducing the number of paired examples shown during training.
+For each image in the test set, generate text conditioned on the image, and measure the number of matching characters (blue line) and the length of text (orange line).
+The figure shows large gains from 0\% to 10\% supervision with much smaller gains as we approach 100\% supervision, suggesting that MMVAEs do not require a lot of paired data to learn a good latent embedding.
+(b) Effect of reducing the number of image/text examples on the character prediction task, and (c) on the length prediction task.
+In (b), the performance is much more sensitive to the number of text samples while in (c), the effect is more balanced between weak image and text supervision.
+TODO: add embedding pca.}
+    \label{fig:multimnist:weaksupervision}
+\end{figure}
+
 \subsection{COCO-Captions}
@@
 VAEs suffer from blurry samples when training on natural images.
-To address this, we replace the DCGAN decoder with a PixelVAE decoder.
+To address this, we replace the DCGAN decoder with a PixelVAE decoder \cite{gulrajani2016pixelvae, oord2016pixel}.
 For the text encoder and decoders, we increase the hidden units to 500.

### 2017-11-18 paper/draft.tex
Finished project update

@@
     \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_on_image_1773_image_sample.png}
+        \caption{}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_on_image_text_417_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_on_text_417_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_on_image_417_image_sample.png}
+        \caption{}
+    \end{subfigure}
+
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_on_image_12_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_on_text_12_image_sample.png}
+        \caption{}
+    \end{subfigure}
+    \begin{subfigure}[b]{.24\linewidth}
+        \centering
+        \includegraphics[width=\linewidth]{multimnist_mmvae_condition_on_image_12_image_sample.png}
+        \caption{}
+    \end{subfigure}
+
     \caption{\emph{MultiMNIST Dataset}: (a) Image samples from $z \sim N(0, 1)$.
@@
 (c) Image samples from $p(z|y=1773)$.
-The quality of the samples from (b) and (c) are almost identical.}
+(d) Image samples from $p(z|x=1773)$.
+(e) Image samples from $p(z|x=417,y=417)$.
+(f) Image samples from $p(z|y=417)$.
+(g) Image samples from $p(z|x=417)$.
+(h) Image samples from $p(z|x=12,y=12)$.
+(i) Image samples from $p(z|y=12)$.
+(j) Image samples from $p(z|x=12)$.
+The quality of samples is consistent across number of digits in the canvas.
+Using image-only, text-only, or both modalities also produce similarly crisp samples.}
     \label{fig:multimnist:samples}
 \end{figure}
+
+\begin{table}[!h]
+    \centering
+    \begin{tabular}{ l | c | c | c | c | c }
+        (a) & 93 & 288 & 403 & 9926 & 21 \\
+        \hline
+        (b) & 1773 & 1773 & 1773 & 1773 & 1773 \\
+        \hline
+        (c) & 1773 & 1773 & 1773 & 1773 & 1773 \\
+        \hline
+        (d) & 1773 & 1773 & 1773 & 1673 & 1773 \\
+        \hline
+        (e) & 417 & 417 & 417 & 417 & 417 \\
+        \hline
+        (f) & 417 & 417 & 417 & 417 & 417 \\
+        \hline
+        (g) & 417 & 417 & 417 & 417 & 417 \\
+        \hline
+        (h) & 21 & 21 & 21 & 21 & 21 \\
+        \hline
+        (i) & 21 & 21 & 21 & 21 & 21 \\
+        \hline
+        (j) & 21 & 21 & 21 & 21 & 21 \\
+        \newline
+    \end{tabular}
+    \caption{\emph{MultiMNIST Dataset}: Generated text samples from MMVAE trained on MultiMNIST.
+(a) Text samples from $z \sim N(0, 1)$.
+(b) Text samples from $p(z|x=1773,y=1773)$.
+(c) Text samples from $p(z|x=1773)$.
+(d) Text samples from $p(z|y=1773)$.
+(e) Text samples from $p(z|x=417,y=417)$.
+(f) Text samples from $p(z|y=417)$.
+(g) Text samples from $p(z|x=417)$.
+(h) Text samples from $p(z|x=12,y=12)$.
+(i) Text samples from $p(z|y=12)$.
+(j) Text samples from $p(z|x=12)$.}
+    \label{table:mnist:samples}
+\end{table}
 

### 2017-11-28 mnist/modal_weak.py
Added modal weak supervision tests to MNIST

@@ -1,0 +1,211 @@
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+
+import os
+import sys
+import shutil
+import random
+import numpy as np
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.autograd import Variable
+from torchvision import datasets, transforms
+
+from model import MultimodalVAE
+from train import AverageMeter
+from train import save_checkpoint, load_checkpoint
+from train import loss_function
+
+
+def train_pipeline(out_dir, weak_perc_m1, weak_perc_m2, n_latents=20, batch_size=128, 
+                   epochs=20, lr=1e-3, log_interval=10, cuda=False):
+    """Pipeline to train and test MultimodalVAE on MNIST dataset. This is 
+    identical to the code in train.py.
+
+    :param out_dir: directory to store trained models
+    :param weak_perc_m1: percent of time to show first modality
+    :param weak_perc_m2: percent of time to show second modality
+    :param n_latents: size of latent variable (default: 20)
+    :param batch_size: number of examples to show at once (default: 128)
+    :param epochs: number of loops over dataset (default: 20)
+    :param lr: learning rate (default: 1e-3)
+    :param log_interval: interval of printing (default: 10)
+    :param cuda: whether to use cuda or not (default: False)
+    """
+    # create loaders for MNIST
+    train_loader = torch.utils.data.DataLoader(
+        datasets.MNIST('./data', train=True, download=True,
+                       transform=transforms.ToTensor()),
+        batch_size=batch_size, shuffle=True)
+    test_loader = torch.utils.data.DataLoader(
+        datasets.MNIST('./data', train=False, download=True,
+                       transform=transforms.ToTensor()),
+        batch_size=batch_size, shuffle=True)
+
+    # load multimodal VAE
+    vae = MultimodalVAE(n_latents=n_latents, use_cuda=cuda)
+    if cuda:
+        vae.cuda()
+
+    optimizer = optim.Adam(vae.parameters(), lr=lr)
+
+
+    def train(epoch):
+        random.seed(42)
+        np.random.seed(42)  # important to have the same seed
+                            # in order to make the same choices for weak supervision
+                            # otherwise, we end up showing different examples over epochs
+        vae.train()
+
[…]

### 2017-11-28 mnist/paired_weak.py

@@ -196,7 +196,7 @@ if __name__ == "__main__":
     args = parser.parse_args()
     args.cuda = args.cuda and torch.cuda.is_available()
 
-    supervision_dir = './trained_models/weak_supervision'
+    supervision_dir = './trained_models/weak_paired_supervision'
     if os.path.isdir(supervision_dir):
         shutil.rmtree(supervision_dir)
     os.makedirs(supervision_dir)

### 2017-11-28 mnist/test_modal_weak.py

@@ -1,0 +1,87 @@
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import operator
+import numpy as np
+import pandas as pd
+
+import torch
+from torchvision import datasets, transforms
+
+from test import test_mnist
+from train import load_checkpoint
+
+
+if __name__ == "__main__":
+    import os
+    import argparse
+    from glob import glob
+
+    import matplotlib
+    matplotlib.use('Agg')
+    import matplotlib.pyplot as plt
+    import seaborn as sns
+
+    parser = argparse.ArgumentParser()
+    parser.add_argument('models_dir', type=str, help='path to output directory of weak.py')
+    parser.add_argument('--cuda', action='store_true', default=False,
+                        help='enables CUDA training')
+    args = parser.parse_args()
+    args.cuda = args.cuda and torch.cuda.is_available()
+
+    x1, x2, y = [], [], []
+
+    for dir_path in glob(os.path.join(args.models_dir, '*')):
+        weak_perc_m1 = float(os.path.basename(dir_path).split('_')[-3])
+        weak_perc_m2 = float(os.path.basename(dir_path).split('_')[-1])
+        loader = torch.utils.data.DataLoader(
+            datasets.MNIST('./data', train=False, download=True,
+                           transform=transforms.ToTensor()),
+            batch_size=128, shuffle=True)
+        vae = load_checkpoint(os.path.join(dir_path, 'model_best.pth.tar'), use_cuda=args.cuda)
+        vae.eval()
+        weak_acc = test_mnist(vae, loader, use_cuda=args.cuda, verbose=False)
+
+        x1.append(weak_perc_m1)
+        x2.append(weak_perc_m2)
+        y.append(weak_acc)
+        print('Got accuracies for %s.' % dir_path)
+
+    assert len(set(x1)) == len(set(x2))
+    percs = sorted(list(set(x1)))
+    n_perc = len(percs)
+
+    x1, x2 = np.array(x1), np.array(x2) 
+    y = np.array(y)
+    data = np.vstack((x1, x2, y)).T.tolist()
+    data = np.array(sorted(data, key=operator.itemgetter(0, 1)))
+    data = data[:, -1].reshape(n_perc, n_perc)
+
+    def save_plot(data, savepath):
+        column_labels = [str(i) for i in percs]
+        row_labels = [str(i) for i in percs]
+    
+        fig, ax = plt.subplots()
+        heatmap = ax.pcolor(data)
+
[…]

### 2017-11-28 mnist/test_paired_weak.py

@@ -54,4 +54,4 @@ if __name__ == "__main__":
     plt.tick_params(axis='both', which='major', labelsize=16)
     plt.tick_params(axis='both', which='minor', labelsize=16)
     plt.tight_layout()
-    plt.savefig('./weak_supervision.png')
+    plt.savefig('./weak_paired_supervision.png')

### 2017-11-28 mnist/modal_weak.py
Bugfixes to modal weak supervision pipeline

@@ -173,8 +173,8 @@ def train_pipeline(out_dir, weak_perc_m1, weak_perc_m2, n_latents=20, batch_size
 if __name__ == "__main__":
     import argparse
     parser = argparse.ArgumentParser()
-    parser.add_argument('--n_latents', type=int, default=100,
-                        help='size of the latent embedding')
+    parser.add_argument('--n_latents', type=int, default=20,
+                        help='size of the latent embedding (default: 20)')
     parser.add_argument('--batch_size', type=int, default=128, metavar='N',
                         help='input batch size for training (default: 128)')
     parser.add_argument('--epochs', type=int, default=20, metavar='N',

### 2017-11-28 mnist/paired_weak.py

@@ -182,7 +182,7 @@ if __name__ == "__main__":
     import argparse
     parser = argparse.ArgumentParser()
     parser.add_argument('--n_latents', type=int, default=20,
-                        help='size of the latent embedding')
+                        help='size of the latent embedding (default: 20)')
     parser.add_argument('--batch_size', type=int, default=128, metavar='N',
                         help='input batch size for training (default: 128)')
     parser.add_argument('--epochs', type=int, default=20, metavar='N',

### 2017-11-28 mnist/modal_weak.py
Bugfix in sign

@@ -48,7 +48,7 @@ def train_pipeline(out_dir, weak_perc_m1, weak_perc_m2, n_latents=20, batch_size
         batch_size=batch_size, shuffle=True)
 
     # load multimodal VAE
-    vae = MultimodalVAE(n_latents=n_latents, use_cuda=cuda)
+    vae = MultimodalVAE(n_latents=n_latents)
     if cuda:
         vae.cuda()
 
@@ -81,7 +81,6 @@ def train_pipeline(out_dir, weak_perc_m1, weak_perc_m2, n_latents=20, batch_size
             # depending on this flip, we decide whether or not to show a modality 
             # versus another one.
             flip = np.random.random()
-
             if flip < weak_perc_m1:
                 recon_image_2, recon_text_2, mu_2, logvar_2 = vae(image=image)
                 loss_2 = loss_function(mu_2, logvar_2, recon_image=recon_image_2, image=image, 
@@ -90,7 +89,7 @@ def train_pipeline(out_dir, weak_perc_m1, weak_perc_m2, n_latents=20, batch_size
                 loss += loss_2
 
             flip = np.random.random()
-            if flip > weak_perc_m2:
+            if flip < weak_perc_m2:
                 recon_image_3, recon_text_3, mu_3, logvar_3 = vae(text=text)
                 loss_3 = loss_function(mu_3, logvar_3, recon_image=recon_image_3, image=image, 
                                        recon_text=recon_text_3, text=text, lambda_xy=0., lambda_yx=1.)

### 2017-11-28 mnist/train.py

@@ -61,6 +61,7 @@ def load_checkpoint(file_path, use_cuda=False):
 def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, text=None,
                   lambda_xy=1., lambda_yx=1.):
     image_BCE, text_BCE = 0, 0
+    batch_size = mu.size(0)
 
     if recon_image is not None and image is not None:
         image_BCE = lambda_xy * F.binary_cross_entropy(recon_image, image.view(-1, 784))
@@ -73,7 +74,7 @@ def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, tex
     # https://arxiv.org/abs/1312.6114
     # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
     KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-    KLD /= args.batch_size * (784 / 3)  # for each pixel
+    KLD /= batch_size * (784 / 3)  # for each pixel
     return image_BCE + text_BCE + KLD
 
 

### 2017-11-28 multimnist/modal_weak.py

@@ -105,7 +105,7 @@ def train_pipeline(out_dir, weak_perc_m1, weak_perc_m2, n_latents=20, batch_size
                 loss += loss_2
 
             flip = np.random.random()
-            if flip > weak_perc_m2:
+            if flip < weak_perc_m2:
                 recon_image_3, recon_text_3, mu_3, logvar_3 = vae(text=text)
                 loss_3 = loss_function(mu_3, logvar_3, recon_image=recon_image_3, image=image, 
                                        recon_text=recon_text_3, text=text, kl_lambda=kl_lambda, 

### 2017-11-28 coco/train.py
Make default image size to generate to be 32

@@ -69,8 +69,8 @@ def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, tex
     image_BCE, text_BCE = 0, 0
     
     if recon_image is not None and image is not None:
-        image_BCE = lambda_xy * F.binary_cross_entropy(recon_image.view(-1, 1 * 64 * 64), 
-                                                       image.view(-1, 1 * 64 * 64))
+        image_BCE = lambda_xy * F.binary_cross_entropy(recon_image.view(-1, 1 * 32 * 32), 
+                                                       image.view(-1, 1 * 32 * 32))
 
     if recon_text is not None and text is not None:
         text_BCE = lambda_yx * F.nll_loss(recon_text.view(-1, recon_text.size(2)), text.view(-1))
@@ -104,11 +104,11 @@ if __name__ == "__main__":
     args = parser.parse_args()
     args.cuda = args.cuda and torch.cuda.is_available()
 
-    transform_train = transforms.Compose([transforms.Scale(64),
-                                          transforms.CenterCrop(64),
+    transform_train = transforms.Compose([transforms.Scale(32),
+                                          transforms.CenterCrop(32),
                                           transforms.ToTensor()])
-    transform_test = transforms.Compose([transforms.Scale(64),
-                                         transforms.CenterCrop(64),
+    transform_test = transforms.Compose([transforms.Scale(32),
+                                         transforms.CenterCrop(32),
                                          transforms.ToTensor()])
     # transformer for text.
     transform_text = text_transformer(deterministic=False)
@@ -253,7 +253,7 @@ if __name__ == "__main__":
                sample = sample.cuda()
 
             image_sample = vae.image_decoder(sample).cpu().data
-            save_image(image_sample.view(64, 3, 64, 64),
+            save_image(image_sample.view(64, 3, 32, 32),
                        './results/sample_image.png')
 
             sample_texts = vae.text_decoder.generate(sample)

### 2017-11-28 coco/train_imageonly.py

@@ -72,11 +72,11 @@ if __name__ == "__main__":
     if not os.path.isdir('./results/image_only'):
         os.makedirs('./results/image_only')
 
-    transform_train = transforms.Compose([transforms.Scale(64),
-                                          transforms.CenterCrop(64),
+    transform_train = transforms.Compose([transforms.Scale(32),
+                                          transforms.CenterCrop(32),
                                           transforms.ToTensor()])
-    transform_test = transforms.Compose([transforms.Scale(64),
-                                         transforms.CenterCrop(64),
+    transform_test = transforms.Compose([transforms.Scale(32),
+                                         transforms.CenterCrop(32),
                                          transforms.ToTensor()])
  
     train_loader = torch.utils.data.DataLoader(
@@ -166,5 +166,5 @@ if __name__ == "__main__":
            sample = sample.cuda()
 
         sample = vae.decode(sample).cpu()
-        save_image(sample.data.view(64, 3, 64, 64),
+        save_image(sample.data.view(64, 3, 32, 32),
                    './results/image_only/sample_image_epoch%d.png' % epoch)

### 2017-11-28 coco/glove.py
Bugfix with GloVe embeddings

@@ -13,7 +13,8 @@ import torchtext.vocab as vocab
 
 
 class GloVe(object):
-    def __init__():
+    def __init__(self):
+        super(GloVe, self).__init__()
         self.glove = vocab.GloVe(name='840B', dim=300)
 
     def get_word(self, word):

### 2017-11-28 coco/model.py
Bugfixes to TextEncoder

@@ -24,8 +24,8 @@ class MultimodalVAE(nn.Module):
         super(MultimodalVAE, self).__init__()
         self.image_encoder = ImageEncoder(n_latents)
         self.image_decoder = ImageDecoder(n_latents)
-        self.text_encoder = TextEncoder(n_latents, n_characters)
-        self.text_decoder = TextDecoder(n_latents, n_characters, use_cuda=use_cuda)
+        self.text_encoder = TextEncoder(n_latents)
+        self.text_decoder = TextDecoder(n_latents, use_cuda=use_cuda)
         self.experts = ProductOfExperts()
 
     def reparametrize(self, mu, logvar):
@@ -60,7 +60,6 @@ class MultimodalVAE(nn.Module):
     def forward(self, image=None, text=None):
         # can't just put nothing
         assert image is not None or text is not None
-        
         if image is not None and text is not None:
             # compute separate gaussians per modality
             image_mu, image_logvar = self.encode_image(image)
@@ -121,8 +120,8 @@ class ImageVAE(nn.Module):
 class TextVAE(nn.Module):
     def __init__(self, n_latents=20, use_cuda=False):
         super(TextVAE, self).__init__()
-        self.encoder = TextEncoder(n_latents, n_characters)
-        self.decoder = TextDecoder(n_latents, n_characters, use_cuda=use_cuda)
+        self.encoder = TextEncoder(n_latents)
+        self.decoder = TextDecoder(n_latents, use_cuda=use_cuda)
         self.n_latents = n_latents
 
     def encode(self, x):

### 2017-11-28 coco/train.py

@@ -118,13 +118,13 @@ if __name__ == "__main__":
         datasets.CocoCaptions('./data/coco/train2014', 
                               './data/coco/annotations/captions_train2014.json',
                               transform=transform_train,
-                              target_transform=coco_char_tensor),
+                              target_transform=transform_text),
         batch_size=args.batch_size, shuffle=True)
     test_loader = torch.utils.data.DataLoader(
         datasets.CocoCaptions('./data/coco/val2014', 
                               './data/coco/annotations/captions_val2014.json',
                               transform=transform_test,
-                              target_transform=coco_char_tensor),
+                              target_transform=transform_text),
         batch_size=args.batch_size, shuffle=True)
 
     # load multimodal VAE

### 2017-11-28 coco/utils.py

@@ -4,6 +4,7 @@ from __future__ import division
 from __future__ import print_function
 from __future__ import absolute_import
 
+import torch
 import random
 from glove import GloVe
 from nltk.tokenize import word_tokenize

### 2017-11-28 coco/glove.py
Bugfixes in glove when word is not recognized

@@ -18,7 +18,9 @@ class GloVe(object):
         self.glove = vocab.GloVe(name='840B', dim=300)
 
     def get_word(self, word):
-        return self.glove.vectors[self.glove.stoi[word]]
+        if word in self.glove.stoi:
+            return self.glove.vectors[self.glove.stoi[word]]
+        return None
 
     def closest(self, vec, n=10):
         """

### 2017-11-28 coco/utils.py

@@ -39,7 +39,9 @@ def text_transformer(deterministic=False):
 
         embeddings = torch.zeros((MAX_WORDS + 2, 300))
         for i, word in enumerate(words):
-            embeddings[i] = glove.get_word(word)
+            glove_vec = glove.get_word(word)
+            if glove_vec is not None:
+                embeddings[i] = glove_vec
 
         return embeddings
 

### 2017-11-28 coco/model.py
Changing images to 32 size affects ImageEncoder

@@ -167,7 +167,7 @@ class ImageEncoder(nn.Module):
             Swish(),
         )
         self.classifier = nn.Sequential(
-            nn.Linear(512 * 4 * 4, 1024),
+            nn.Linear(512 * 2 * 2, 1024),
             Swish(),
             nn.Dropout(p=0.1),
             nn.Linear(1024, 256),
@@ -180,7 +180,7 @@ class ImageEncoder(nn.Module):
     def forward(self, x):
         n_latents = self.n_latents
         x = self.features(x)
-        x = x.view(-1, 512 * 4 * 4)
+        x = x.view(-1, 512 * 2 * 2)
         x = self.classifier(x)
         return x[:, :n_latents], x[:, n_latents:]
 

### 2017-11-28 coco/model.py
Bugfixes

@@ -189,7 +189,7 @@ class ImageDecoder(nn.Module):
     def __init__(self, n_latents):
         super(ImageDecoder, self).__init__()
         self.upsample = nn.Sequential(
-            nn.Linear(n_latents, 512 * 4 * 4),
+            nn.Linear(n_latents, 512 * 2 * 2),
             Swish(),
         )
         self.hallucinate = nn.Sequential(
@@ -208,7 +208,7 @@ class ImageDecoder(nn.Module):
     def forward(self, z):
         # the input will be a vector of size |n_latents|
         z = self.upsample(z)
-        z = z.view(-1, 512, 4, 4)
+        z = z.view(-1, 512, 2, 2)
         z = self.hallucinate(z)
         return F.sigmoid(z)
 
@@ -250,15 +250,16 @@ class TextDecoder(nn.Module):
                         (default: 300)
     :param use_cuda: whether to use cuda tensors
     """
-    def __init__(self, n_latents, n_embedding=300, use_cuda=False):
+    def __init__(self, n_latents, n_embedding=300, n_hiddens=200,  use_cuda=False):
         super(TextDecoder, self).__init__()
-        self.z2h = nn.Linear(n_latents, 200)
-        self.gru = nn.GRU(n_embedding + n_latents, 200, 2, dropout=0.1)
-        self.h2o = nn.Linear(n_embedding + n_latents, n_embedding)
+        self.z2h = nn.Linear(n_latents, n_hiddens)
+        self.gru = nn.GRU(n_embedding + n_latents, n_hiddens, 2, dropout=0.1)
+        self.h2o = nn.Linear(n_hiddens + n_latents, n_embedding)
         self.glove = GloVe()
         self.use_cuda = use_cuda
         self.n_latents = n_latents
         self.n_embedding = n_embedding
+        self.n_hiddens = n_hiddens
 
     def forward(self, z):
         n_latents = self.n_latents
@@ -279,10 +280,10 @@ class TextDecoder(nn.Module):
         # look through n_steps and generate characters
         for i in xrange(MAX_WORDS):
             w_out, h = self.step(i, z, w_in, h)
-            words[:, i] = w_out
+            sentence[:, i] = w_out
             w_in = w_out
 
-        return words  # (batch_size, seq_len, ...)
+        return sentence  # (batch_size, seq_len, ...)
 
     def generate(self, z):
         words = self.forward(z)  # shape is (batch_size, seq_len, n_embedding)

### 2017-11-28 coco/train.py

@@ -69,8 +69,8 @@ def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, tex
     image_BCE, text_BCE = 0, 0
     
     if recon_image is not None and image is not None:
-        image_BCE = lambda_xy * F.binary_cross_entropy(recon_image.view(-1, 1 * 32 * 32), 
-                                                       image.view(-1, 1 * 32 * 32))
+        image_BCE = lambda_xy * F.binary_cross_entropy(recon_image.view(-1, 3 * 32 * 32), 
+                                                       image.view(-1, 3 * 32 * 32))
 
     if recon_text is not None and text is not None:
         text_BCE = lambda_yx * F.nll_loss(recon_text.view(-1, recon_text.size(2)), text.view(-1))

### 2017-11-28 coco/train.py
Switched Text Loss to MSE

@@ -73,7 +73,7 @@ def loss_function(mu, logvar, recon_image=None, image=None, recon_text=None, tex
                                                        image.view(-1, 3 * 32 * 32))
 
     if recon_text is not None and text is not None:
-        text_BCE = lambda_yx * F.nll_loss(recon_text.view(-1, recon_text.size(2)), text.view(-1))
+        text_BCE = lambda_yx * F.mse_loss(recon_text, text)
 
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014

### 2017-11-28 coco/utils.py

@@ -12,6 +12,7 @@ from nltk.tokenize import word_tokenize
 MAX_WORDS = 100  # max number of words in a sentence
 SOS = '<s>'
 EOS = '</s>'
+MAX_WORDS += 2
 
 
 def text_transformer(deterministic=False):
@@ -37,7 +38,7 @@ def text_transformer(deterministic=False):
             words = words[:MAX_WORDS]
         words = [SOS] + words + [EOS]
 
-        embeddings = torch.zeros((MAX_WORDS + 2, 300))
+        embeddings = torch.zeros((MAX_WORDS, 300))
         for i, word in enumerate(words):
             glove_vec = glove.get_word(word)
             if glove_vec is not None:

### 2017-11-28 mnist/test_modal_weak.py
Reversed labels of text and image

@@ -78,8 +78,8 @@ if __name__ == "__main__":
         ax.set_yticklabels(column_labels, minor=False)
         
         plt.colorbar(heatmap)
-        plt.xlabel('% Image Examples', fontsize=18)
-        plt.ylabel('% Text Examples', fontsize=18)
+        plt.xlabel('% Text Examples', fontsize=18)
+        plt.ylabel('% Image Examples', fontsize=18)
 
         plt.tight_layout()
         plt.savefig(savepath)

### 2017-11-28 multimnist/test_modal_weak.py

@@ -85,8 +85,8 @@ if __name__ == "__main__":
         ax.set_yticklabels(column_labels, minor=False)
         
         plt.colorbar(heatmap)
-        plt.xlabel('% Image Examples', fontsize=18)
-        plt.ylabel('% Text Examples', fontsize=18)
+        plt.xlabel('% Text Examples', fontsize=18)
+        plt.ylabel('% Image Examples', fontsize=18)
 
         plt.tight_layout()
         plt.savefig(savepath)

### 2017-11-28 coco/glove.py
Don't do GloVe lookups every iter -- just save the raw GloVe vectors outputted

@@ -9,6 +9,7 @@ from __future__ import print_function
 from __future__ import absolute_import
 
 import torch
+from torch.autograd import Variable
 import torchtext.vocab as vocab
 
 
@@ -47,11 +48,16 @@ class GloVe(object):
         batch_size = len(vec_batch)
         dict_size = len(self.glove.itos)
         all_dists = torch.zeros((batch_size, dict_size))
+        use_cuda = vec_batch.is_cuda
 
         for iw, word in enumerate(self.glove.itos):
             for iv, vec in enumerate(vec_batch):
-                dist = torch.dist(vec, self.get_word(word ))
-                all_dists[iv, iw] = dist
+                word_vec = self.get_word(word)
+                if use_cuda:
+                    word_vec = word_vec.cuda()
+                word_vec = Variable(word_vec)
+                dist = torch.dist(vec, word_vec)
+                all_dists[iv, iw] = dist.data[0]
 
         top_dist = torch.zeros(batch_size)
         for ix in xrange(batch_size):

### 2017-11-28 coco/model.py

@@ -303,6 +303,9 @@ class TextDecoder(nn.Module):
 
         return reshape
 
+    def generate_vector(self, z):
+        return self.forward(z)
+
     def step(self, ix, z, w_in, h):
         # w_in is a (batch, n_embedding)
         w_in = torch.cat((w_in, z), dim=1)  # n_embedding + n_latents

### 2017-11-28 coco/train.py

@@ -247,6 +247,8 @@ if __name__ == "__main__":
             'optimizer' : optimizer.state_dict(),
         }, is_best, folder='./trained_models')   
 
+        is_best = True
+
         if is_best:
             sample = Variable(torch.randn(64, args.n_latents))
             if args.cuda:
@@ -256,6 +258,5 @@ if __name__ == "__main__":
             save_image(image_sample.view(64, 3, 32, 32),
                        './results/sample_image.png')
 
-            sample_texts = vae.text_decoder.generate(sample)
-            with open('./results/sample_text.txt', 'w') as fp:
-                fp.writelines(sample_texts)
+            sample_texts = vae.text_decoder.generate_vector(sample).data
+            torch.save(sample_texts, './results/sample_text_vector.pt')

### 2017-11-29 mnist/model.py
Added InfoVAE implementation

@@ -232,6 +232,86 @@ class VAE(nn.Module):
         return self.decode(z), mu, logvar
 
 
+# -- Begin InfoVAE section --
+
+
+class InfoVAE(nn.Module):
+    def __init__(self, n_latents=20):
+        super(InfoVAE, self).__init__()
+        self.n_latents = n_latents
+        self.encoder_conv = nn.Sequential(
+            nn.Conv2d(1, 64, 4, 2),
+            nn.LeakyReLU(0.1, inplace=True),
+            nn.Conv2d(64, 128, 4, 2),
+            nn.LeakyReLU(0.1, inplace=True),
+        )
+        self.encoder_fc = nn.Sequential(
+            nn.Linear(128 * 5 * 5, 1024),
+            nn.LeakyReLU(0.1, inplace=True),
+            nn.Linear(1024, n_latents * 2)   
+        )
+        self.decoder_fc = nn.Sequential(
+            nn.Linear(n_latents, 1024),
+            nn.ReLU(True),
+            nn.Linear(1024, 128 * 7 * 7),
+            nn.ReLU(True),
+        )
+        self.decoder_conv = nn.Sequential(
+            nn.ConvTranspose2d(128, 64, 4, 2),
+            nn.ReLU(True),
+            nn.ConvTranspose2d(64, 1, 4, 2),
+        )
+
+    def encode(self, x):
+        x = self.encoder_conv(x)
+        x = x.view(x.size(0), -1)
+        x = self.encoder_fc(x)
+        n_latents = self.n_latents
+        return x[:, :n_latents], x[:, n_latents:]
+
+    def reparametrize(self, mu, logvar):
+        if self.training:
+            std = logvar.mul(0.5).exp_()
+            eps = Variable(std.data.new(std.size()).normal_())
+            return eps.mul(std).add_(mu)
+        else:
+            return mu
+
+    def decode(self, z):
+        z = self.decoder_fc(z)
+        z = z.view(-1, 128, 7, 7)
+        z = self.decoder_conv(z)
+        return F.sigmoid(z)
+
+    def forward(self, x):
+        mu, logvar = self.encode(x)
+        z = self.reparameterize(mu, logvar)
+        return self.decode(z), mu, logvar
+
+
+def compute_kernel(x, y):
+    """Apply Gaussian kernel to the i-th vector of x and j-th vector of y.
+
[…]

### 2017-11-29 mnist/train_infovae.py

@@ -1,0 +1,147 @@
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import os
+import sys
+import shutil
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.autograd import Variable
+from torchvision import datasets, transforms
+
+from model import InfoVAE
+from train import AverageMeter
+
+
+def save_checkpoint(state, is_best, folder='./', filename='checkpoint.pth.tar'):
+    torch.save(state, os.path.join(folder, filename))
+    if is_best:
+        shutil.copyfile(os.path.join(folder, filename),
+                        os.path.join(folder, 'model_best.pth.tar'))
+
+
+def load_checkpoint(file_path, use_cuda=False):
+    """Return EmbedNet instance"""
+    if use_cuda:
+        checkpoint = torch.load(file_path)
+    else:
+        checkpoint = torch.load(file_path,
+                                map_location=lambda storage, location: storage)
+    n_latents = checkpoint['n_latents']
+    vae = InfoVAE(n_latents=n_latents)
+    vae.load_state_dict(checkpoint['state_dict'])
+    
+    if use_cuda:
+        vae.cuda()
+
+    return vae
+
+
+def loss_function(recon_x, x, mu, logvar):
+    BCE = F.binary_cross_entropy(recon_x, x)
+    # see Appendix B from VAE paper:
+    # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
+    # https://arxiv.org/abs/1312.6114
+    # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
+    KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
+    KLD /= batch_size * 784
+    return BCE + KLD
+
+
+if __name__ == "__main__":
+    import argparse
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--n_latents', type=int, default=20,
+                        help='size of the latent embedding')
+    parser.add_argument('--batch_size', type=int, default=128, metavar='N',
+                        help='input batch size for training (default: 128)')
+    parser.add_argument('--epochs', type=int, default=20, metavar='N',
+                        help='number of epochs to train (default: 20)')
+    parser.add_argument('--lr', type=float, default=1e-3, metavar='LR',
+                        help='learning rate (default: 1e-3)')
+    parser.add_argument('--log_interval', type=int, default=10, metavar='N',
[…]

### 2017-11-29 coco/train.py
Added bugfixes to get InfoVAE working

@@ -256,7 +256,7 @@ if __name__ == "__main__":
 
             image_sample = vae.image_decoder(sample).cpu().data
             save_image(image_sample.view(64, 3, 32, 32),
-                       './results/sample_image.png')
+                       './results/sample_image_epoch%d.png' % epoch)
 
             sample_texts = vae.text_decoder.generate_vector(sample).data
             torch.save(sample_texts, './results/sample_text_vector.pt')

### 2017-11-29 mnist/model.py

@@ -253,11 +253,11 @@ class InfoVAE(nn.Module):
         self.decoder_fc = nn.Sequential(
             nn.Linear(n_latents, 1024),
             nn.ReLU(True),
-            nn.Linear(1024, 128 * 7 * 7),
+            nn.Linear(1024, 128 * 5 * 5),
             nn.ReLU(True),
         )
         self.decoder_conv = nn.Sequential(
-            nn.ConvTranspose2d(128, 64, 4, 2),
+            nn.ConvTranspose2d(128, 64, 5, 2),
             nn.ReLU(True),
             nn.ConvTranspose2d(64, 1, 4, 2),
         )
@@ -279,13 +279,13 @@ class InfoVAE(nn.Module):
 
     def decode(self, z):
         z = self.decoder_fc(z)
-        z = z.view(-1, 128, 7, 7)
+        z = z.view(-1, 128, 5, 5)
         z = self.decoder_conv(z)
         return F.sigmoid(z)
 
     def forward(self, x):
         mu, logvar = self.encode(x)
-        z = self.reparameterize(mu, logvar)
+        z = self.reparametrize(mu, logvar)
         return self.decode(z), mu, logvar
 
 

### 2017-11-29 mnist/train_infovae.py

@@ -42,6 +42,7 @@ def load_checkpoint(file_path, use_cuda=False):
 
 
 def loss_function(recon_x, x, mu, logvar):
+    batch_size = x.size(0)
     BCE = F.binary_cross_entropy(recon_x, x)
     # see Appendix B from VAE paper:
     # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
@@ -98,16 +99,16 @@ if __name__ == "__main__":
             data = Variable(data)
             optimizer.zero_grad()
             
-            recon_x, recon_x, mu, logvar = vae(data)
-            loss = loss_function(recon_x, x, mu, logvar)
-            loss_meter.update(loss.data[0], len(x))
+            recon_data, mu, logvar = vae(data)
+            loss = loss_function(recon_data, data, mu, logvar)
+            loss_meter.update(loss.data[0], len(data))
 
             loss.backward()
             optimizer.step()
 
             if batch_idx % args.log_interval == 0:
                 print('Train Epoch: {} [{}/{} ({:.0f}%)]\tLoss: {:.6f}'.format(
-                    epoch, batch_idx * len(image), len(train_loader.dataset),
+                    epoch, batch_idx * len(data), len(train_loader.dataset),
                     100. * batch_idx / len(train_loader), loss_meter.avg))
 
         print('====> Epoch: {}\tLoss: {:.4f}'.format(epoch, loss_meter.avg))
@@ -122,8 +123,8 @@ if __name__ == "__main__":
                 data = data.cuda()
             data = Variable(data)
             
-            recon_x, recon_x, mu, logvar = vae(data)
-            loss = loss_function(recon_x, x, mu, logvar)                
+            recon_data, mu, logvar = vae(data)
+            loss = loss_function(recon_data, data, mu, logvar)                
             test_loss += loss.data[0]
 
         test_loss /= len(test_loader)

### 2017-11-29 mnist/train_infovae.py
Added sampling every iter

@@ -145,3 +145,11 @@ if __name__ == "__main__":
             'n_latents': args.n_latents,
             'optimizer' : optimizer.state_dict(),
         }, is_best, folder='./trained_models/infovae')     
+
+        sample = Variable(torch.randn(64, args.n_latents))
+        if args.cuda:
+           sample = sample.cuda()
+
+        sample = vae.decode(sample).cpu().data
+        save_image(sample.view(64, 1, 28, 28),
+                   './results/sample_epoch%d.png' % epoch)

### 2017-11-29 mnist/train_infovae.py
InfoVAE is working on MNIST

@@ -12,6 +12,7 @@ import torch.optim as optim
 import torch.nn.functional as F
 from torch.autograd import Variable
 from torchvision import datasets, transforms
+from torchvision.utils import save_image
 
 from model import InfoVAE
 from train import AverageMeter
@@ -153,4 +154,4 @@ if __name__ == "__main__":
 
         sample = vae.decode(sample).cpu().data
         save_image(sample.view(64, 1, 28, 28),
-                   './results/sample_epoch%d.png' % epoch)
+                   './results/infovae/sample_epoch%d.png' % epoch)

### 2017-11-29 mnist/model.py
Switch loss to MMD

@@ -286,7 +286,7 @@ class InfoVAE(nn.Module):
     def forward(self, x):
         mu, logvar = self.encode(x)
         z = self.reparametrize(mu, logvar)
-        return self.decode(z), mu, logvar
+        return self.decode(z), z
 
 
 def compute_kernel(x, y):

### 2017-11-29 mnist/train_infovae.py

@@ -15,6 +15,7 @@ from torchvision import datasets, transforms
 from torchvision.utils import save_image
 
 from model import InfoVAE
+from model import compute_mmd
 from train import AverageMeter
 
 
@@ -42,16 +43,15 @@ def load_checkpoint(file_path, use_cuda=False):
     return vae
 
 
-def loss_function(recon_x, x, mu, logvar):
-    batch_size = x.size(0)
+def loss_function(recon_x, x, z):
+    batch_size = z.size(0)
     BCE = F.binary_cross_entropy(recon_x, x)
-    # see Appendix B from VAE paper:
-    # Kingma and Welling. Auto-Encoding Variational Bayes. ICLR, 2014
-    # https://arxiv.org/abs/1312.6114
-    # 0.5 * sum(1 + log(sigma^2) - mu^2 - sigma^2)
-    KLD = -0.5 * torch.sum(1 + logvar - mu.pow(2) - logvar.exp())
-    KLD /= batch_size * 784
-    return BCE + KLD
+    # Compare the generated z with true samples from a standard Gaussian
+    # and compute their MMD distance
+    true_samples = torch.normal(torch.zeros(batch_size, args.n_latents),
+                                torch.ones(batch_size, args.n_latents))
+    MMD = compute_mmd(true_samples, z)
+    return BCE + MMD
 
 
 if __name__ == "__main__":
@@ -100,8 +100,8 @@ if __name__ == "__main__":
             data = Variable(data)
             optimizer.zero_grad()
             
-            recon_data, mu, logvar = vae(data)
-            loss = loss_function(recon_data, data, mu, logvar)
+            recon_data, z = vae(data)
+            loss = loss_function(recon_data, data, z)
             loss_meter.update(loss.data[0], len(data))
 
             loss.backward()
@@ -124,8 +124,8 @@ if __name__ == "__main__":
                 data = data.cuda()
             data = Variable(data)
             
-            recon_data, mu, logvar = vae(data)
-            loss = loss_function(recon_data, data, mu, logvar)                
+            recon_data, z = vae(data)
+            loss = loss_function(recon_data, data, z)                
             test_loss += loss.data[0]
 
         test_loss /= len(test_loader)

### 2017-11-29 coco/model.py
Added InfoVAE for COCO -- let's see if it makes sharper images

@@ -353,6 +353,58 @@ class Swish(nn.Module):
         return x * F.sigmoid(x)
 
 
+# ---- Start InfoVAE Section ----
+
+class InfoVAE(nn.Module):
+    def __init__(self, n_latents=20):
+        super(InfoVAE, self).__init__()
+        self.encoder = ImageEncoder(n_latents)
+        self.decoder = ImageDecoder(n_latents)
+        self.n_latents = n_latents
+
+    def encode(self, x):
+        return self.encoder(x)
+
+    def reparametrize(self, mu, logvar):
+        if self.training:
+            std = logvar.mul(0.5).exp_()
+            eps = Variable(std.data.new(std.size()).normal_())
+            return eps.mul(std).add_(mu)
+        else:  # return mean during inference
+            return mu
+
+    def decode(self, z):
+        return self.decoder(z)
+
+    def forward(self, x):
+        mu, logvar = self.encode(x)
+        z = self.reparametrize(mu, logvar)
+        return self.decode(z), z
+
+
+def compute_kernel(x, y):
+    """Apply Gaussian kernel to the i-th vector of x and j-th vector of y.
+
+    :param x: torch.Tensor (x_size, dim)
+    :param y: torch.Tensor (y_size, dim)
+    """
+    x_size, y_size, dim = x.size(0), y.size(0), x.size(1)
+    tiled_x = x.unsqueeze(1).expand(x_size, y_size, dim)
+    tiled_y = y.unsqueeze(0).expand(x_size, y_size, dim)
+    return torch.exp(-torch.mean(torch.pow(tiled_x - tiled_y, 2), dim=2) / dim)
+
+
+def compute_mmd(x, y):
+    """Compute maximum mean discrepancy."""
+    x_kernel = compute_kernel(x, x)
+    y_kernel = compute_kernel(y, y)
+    xy_kernel = compute_kernel(x, y)
+    return torch.mean(x_kernel) + torch.mean(y_kernel) - 2 * torch.mean(xy_kernel)
+
+
+# ---- Start PixelCNN Section ----
+
+
 class PixelCNN(nn.Module):
     """Vanilla PixelCNN from first Van de Oord paper."""
     def __init__(self, n_blocks=15, data_channels=1, hid_dims=128, out_dims=256):

### 2017-11-29 coco/train_infovae.py

@@ -1,0 +1,169 @@
+from __future__ import division
+from __future__ import print_function
+from __future__ import absolute_import
+
+import os
+import shutil
+import sys
+import numpy as np
+
+import torch
+import torch.nn as nn
+import torch.optim as optim
+import torch.nn.functional as F
+from torch.autograd import Variable
+from torchvision import datasets, transforms
+from torchvision.utils import save_image
+
+from model import InfoVAE
+from model import compute_mmd
+from train import AverageMeter
+
+
+def save_checkpoint(state, is_best, folder='./', filename='checkpoint.pth.tar'):
+    torch.save(state, os.path.join(folder, filename))
+    if is_best:
+        shutil.copyfile(os.path.join(folder, filename),
+                        os.path.join(folder, 'model_best.pth.tar'))
+
+
+def load_checkpoint(file_path, use_cuda=False):
+    """Return EmbedNet instance"""
+    if use_cuda:
+        checkpoint = torch.load(file_path)
+    else:
+        checkpoint = torch.load(file_path,
+                                map_location=lambda storage, location: storage)
+
+    vae = InfoVAE(n_latents=checkpoint['n_latents'])
+    vae.load_state_dict(checkpoint['state_dict'])
+    
+    return vae
+
+
+def loss_function(recon_x, x, z):
+    batch_size = z.size(0)
+    BCE = F.binary_cross_entropy(recon_x, x)
+    # Compare the generated z with true samples from a standard Gaussian
+    # and compute their MMD distance
+    true_samples = torch.normal(torch.zeros(batch_size, args.n_latents),
+                                torch.ones(batch_size, args.n_latents))
+    MMD = compute_mmd(true_samples, z)
+    return BCE + MMD
+
+
+if __name__ == "__main__":
+    import argparse
+    parser = argparse.ArgumentParser()
+    parser.add_argument('--n_latents', type=int, default=20,
+                        help='size of the latent embedding')
+    parser.add_argument('--batch_size', type=int, default=128, metavar='N',
+                        help='input batch size for training (default: 128)')
+    parser.add_argument('--epochs', type=int, default=10, metavar='N',
+                        help='number of epochs to train (default: 10)')
+    parser.add_argument('--lr', type=float, default=1e-4, metavar='LR',
+                        help='learning rate (default: 1e-4)')
+    parser.add_argument('--log_interval', type=int, default=10, metavar='N',
[…]

### 2017-11-29 coco/train_infovae.py
Bugfixes with infoVAE

@@ -48,6 +48,9 @@ def loss_function(recon_x, x, z):
     # and compute their MMD distance
     true_samples = torch.normal(torch.zeros(batch_size, args.n_latents),
                                 torch.ones(batch_size, args.n_latents))
+    if z.is_cuda:
+        true_samples = true_samples.cuda()
+        true_samples = Variable(true_samples)
     MMD = compute_mmd(true_samples, z)
     return BCE + MMD
 
@@ -55,8 +58,8 @@ def loss_function(recon_x, x, z):
 if __name__ == "__main__":
     import argparse
     parser = argparse.ArgumentParser()
-    parser.add_argument('--n_latents', type=int, default=20,
-                        help='size of the latent embedding')
+    parser.add_argument('--n_latents', type=int, default=100,
+                        help='size of the latent embedding (default: 100)')
     parser.add_argument('--batch_size', type=int, default=128, metavar='N',
                         help='input batch size for training (default: 128)')
     parser.add_argument('--epochs', type=int, default=10, metavar='N',

### 2017-11-29 mnist/train_infovae.py

@@ -50,6 +50,9 @@ def loss_function(recon_x, x, z):
     # and compute their MMD distance
     true_samples = torch.normal(torch.zeros(batch_size, args.n_latents),
                                 torch.ones(batch_size, args.n_latents))
+    if z.is_cuda:
+        true_samples = true_samples.cuda()
+        true_samples = Variable(true_samples)
     MMD = compute_mmd(true_samples, z)
     return BCE + MMD
 

### 2017-11-29 mnist/model.py
Forgot to turn bias off in net

@@ -240,26 +240,26 @@ class InfoVAE(nn.Module):
         super(InfoVAE, self).__init__()
         self.n_latents = n_latents
         self.encoder_conv = nn.Sequential(
-            nn.Conv2d(1, 64, 4, 2),
+            nn.Conv2d(1, 64, 4, 2, 1, bias=False),
             nn.LeakyReLU(0.1, inplace=True),
-            nn.Conv2d(64, 128, 4, 2),
+            nn.Conv2d(64, 128, 4, 2, 1, bias=False),
             nn.LeakyReLU(0.1, inplace=True),
         )
         self.encoder_fc = nn.Sequential(
-            nn.Linear(128 * 5 * 5, 1024),
+            nn.Linear(128 * 7 * 7, 1024),
             nn.LeakyReLU(0.1, inplace=True),
             nn.Linear(1024, n_latents * 2)   
         )
         self.decoder_fc = nn.Sequential(
             nn.Linear(n_latents, 1024),
             nn.ReLU(True),
-            nn.Linear(1024, 128 * 5 * 5),
+            nn.Linear(1024, 128 * 7 * 7),
             nn.ReLU(True),
         )
         self.decoder_conv = nn.Sequential(
-            nn.ConvTranspose2d(128, 64, 5, 2),
+            nn.ConvTranspose2d(128, 64, 4, 2, 1, bias=False),
             nn.ReLU(True),
-            nn.ConvTranspose2d(64, 1, 4, 2),
+            nn.ConvTranspose2d(64, 1, 4, 2, 1, bias=False),
         )
 
     def encode(self, x):
@@ -279,7 +279,7 @@ class InfoVAE(nn.Module):
 
     def decode(self, z):
         z = self.decoder_fc(z)
-        z = z.view(-1, 128, 5, 5)
+        z = z.view(-1, 128, 7, 7)
         z = self.decoder_conv(z)
         return F.sigmoid(z)
 

### 2017-11-29 coco/train_infovae.py
Directly output z instead of mu and logvar ?

@@ -120,7 +120,6 @@ if __name__ == "__main__":
                 data = data.cuda()
             optimizer.zero_grad()
             recon_data, z = vae(data)
-            # watch out for logvar -- could explode if learning rate is too high.  
             loss = loss_function(recon_data, data, z)
             loss_meter.update(loss.data[0], len(data))            
             loss.backward()

### 2017-11-29 mnist/model.py

@@ -248,7 +248,7 @@ class InfoVAE(nn.Module):
         self.encoder_fc = nn.Sequential(
             nn.Linear(128 * 7 * 7, 1024),
             nn.LeakyReLU(0.1, inplace=True),
-            nn.Linear(1024, n_latents * 2)   
+            nn.Linear(1024, n_latents)   
         )
         self.decoder_fc = nn.Sequential(
             nn.Linear(n_latents, 1024),
@@ -266,16 +266,7 @@ class InfoVAE(nn.Module):
         x = self.encoder_conv(x)
         x = x.view(x.size(0), -1)
         x = self.encoder_fc(x)
-        n_latents = self.n_latents
-        return x[:, :n_latents], x[:, n_latents:]
-
-    def reparametrize(self, mu, logvar):
-        if self.training:
-            std = logvar.mul(0.5).exp_()
-            eps = Variable(std.data.new(std.size()).normal_())
-            return eps.mul(std).add_(mu)
-        else:
-            return mu
+        return x
 
     def decode(self, z):
         z = self.decoder_fc(z)
@@ -284,8 +275,7 @@ class InfoVAE(nn.Module):
         return F.sigmoid(z)
 
     def forward(self, x):
-        mu, logvar = self.encode(x)
-        z = self.reparametrize(mu, logvar)
+        z = self.encode(x)
         return self.decode(z), z
 
 

### 2017-11-29 mnist/model.py
Use MSE instead of BCE

@@ -279,26 +279,6 @@ class InfoVAE(nn.Module):
         return self.decode(z), z
 
 
-def compute_kernel(x, y):
-    """Apply Gaussian kernel to the i-th vector of x and j-th vector of y.
-
-    :param x: torch.Tensor (x_size, dim)
-    :param y: torch.Tensor (y_size, dim)
-    """
-    x_size, y_size, dim = x.size(0), y.size(0), x.size(1)
-    tiled_x = x.unsqueeze(1).expand(x_size, y_size, dim)
-    tiled_y = y.unsqueeze(0).expand(x_size, y_size, dim)
-    return torch.exp(-torch.mean(torch.pow(tiled_x - tiled_y, 2), dim=2) / dim)
-
-
-def compute_mmd(x, y):
-    """Compute maximum mean discrepancy."""
-    x_kernel = compute_kernel(x, x)
-    y_kernel = compute_kernel(y, y)
-    xy_kernel = compute_kernel(x, y)
-    return torch.mean(x_kernel) + torch.mean(y_kernel) - 2 * torch.mean(xy_kernel)
-
-
 # -- Begin PixelCNN Section -- 
 
 

### 2017-11-29 mnist/train_infovae.py

@@ -15,7 +15,6 @@ from torchvision import datasets, transforms
 from torchvision.utils import save_image
 
 from model import InfoVAE
-from model import compute_mmd
 from train import AverageMeter
 
 
@@ -45,7 +44,7 @@ def load_checkpoint(file_path, use_cuda=False):
 
 def loss_function(recon_x, x, z):
     batch_size = z.size(0)
-    BCE = F.binary_cross_entropy(recon_x, x)
+    NLL = torch.mean(torch.pow(recon_x - x, 2))
     # Compare the generated z with true samples from a standard Gaussian
     # and compute their MMD distance
     true_samples = torch.normal(torch.zeros(batch_size, args.n_latents),
@@ -54,7 +53,27 @@ def loss_function(recon_x, x, z):
         true_samples = true_samples.cuda()
         true_samples = Variable(true_samples)
     MMD = compute_mmd(true_samples, z)
-    return BCE + MMD
+    return NLL + MMD
+
+
+def compute_kernel(x, y):
+    """Apply Gaussian kernel to the i-th vector of x and j-th vector of y.
+
+    :param x: torch.Tensor (x_size, dim)
+    :param y: torch.Tensor (y_size, dim)
+    """
+    x_size, y_size, dim = x.size(0), y.size(0), x.size(1)
+    tiled_x = x.unsqueeze(1).expand(x_size, y_size, dim)
+    tiled_y = y.unsqueeze(0).expand(x_size, y_size, dim)
+    return torch.exp(-torch.mean(torch.pow(tiled_x - tiled_y, 2), dim=2) / float(dim))
+
+
+def compute_mmd(x, y):
+    """Compute maximum mean discrepancy."""
+    x_kernel = compute_kernel(x, x)
+    y_kernel = compute_kernel(y, y)
+    xy_kernel = compute_kernel(x, y)
+    return torch.mean(x_kernel) + torch.mean(y_kernel) - 2 * torch.mean(xy_kernel)
 
 
 if __name__ == "__main__":
